\section{Artifact and Source Table Manifest}

The main paper is backed by the source tables copied into
\texttt{neuripsnewpaper/source\_tables/} and by the imported full Self-RAG
artifact files listed below.

\begin{table}[H]
\centering
\small
\caption{Primary source artifacts used in this draft.}
\begin{tabular}{ll}
\toprule
File & Used for \\
\midrule
\texttt{paired\_wilcoxon.csv} & Matched-similarity causal intervention \\
\texttt{headline\_table.csv} & Scaled SQuAD/Mistral headline cell \\
\texttt{paired\_contrasts.csv} & Per-seed paired tests for Fix 2 \\
\texttt{table1\_multimetric.csv} & Three-metric faithfulness table \\
\texttt{metric\_correlations.csv} & Metric agreement analysis \\
\texttt{tau\_transfer\_matrix.csv} & Cross-dataset threshold transfer \\
\texttt{generalization\_flags.csv} & Dataset-specific tuning flags \\
\texttt{slope\_response.csv} & Coherence-preserving noise slopes \\
\texttt{h2h\_summary\_full\_selfrag.csv} & Full baseline comparison with Self-RAG \\
\texttt{data/revision/fix\_06/per\_query\_full\_selfrag.csv} & Full Self-RAG per-query artifact \\
\texttt{results/revision/fix\_06/h2h\_summary\_full\_selfrag.csv} & Imported full Self-RAG summary \\
\texttt{raptor\_full\_table.csv} & RAPTOR latency and indexing cost \\
\texttt{partial\_correlations.csv} & Self-confidence association audit \\
\texttt{source\_tables/CHECKSUMS.md} & Checksums for imported source tables \\
\bottomrule
\end{tabular}
\end{table}

\section{Human Calibration Plan}

No completed two-rater human labels are reported in the main paper. Two staged
files exist:
\begin{itemize}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
  \item \path{data/revision/fix_03/human_eval_template.jsonl}: 99 examples
  with blank rater fields.
  \item \path{results/human_eval/samples.csv}: an older 90-example sampling
  sheet with blank human fields.
\end{itemize}
The planned calibration uses two independent raters, a three-way label scheme
(supported, partially supported, unsupported), Cohen's $\kappa$ for inter-rater
agreement, and Spearman correlation between adjudicated labels and each
automatic score: DeBERTa, the second NLI scorer, and the local RAGAS-style
judge.

\section{Supplementary Submission Notes}

\begin{enumerate}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
  \item Fill the 99-item human-evaluation template and report rater agreement.
  \item Add the official \texttt{neurips\_2026.sty} file before submission.
  The draft source already loads it automatically if present; the current
  geometry block is a local fallback only.
  \item Keep the main paper and supplement separate for submission unless the
  final NeurIPS instructions require a single combined PDF.
  \item Replace anonymous metadata with camera-ready metadata only after review.
\end{enumerate}
