\section{Controlled Experimental Design}

The experiments are a controlled falsification suite. Each test isolates a
claim that can look convincing in a conventional aggregate RAG table, then
asks whether that claim survives when the relevant confounder is separated.

\paragraph{Pre-registration.}
Sample sizes, seeds, primary statistics, and decision rules were frozen before
execution. Paired continuous contrasts use Wilcoxon signed-rank tests and
Cohen's $d_z$. Means use 10,000-resample bootstrap confidence intervals.
Hallucination rates use Wilson intervals. HCPC thresholds remain frozen except
in the threshold-transfer experiment, where tuning and evaluation datasets are
explicitly separated. No paid API calls are used in the revision results.

\paragraph{Test 1: separating coherence from local relevance.}
For each SQuAD query, we retrieve the top 20 passages and search over
3-passage subsets. We choose one HIGH-CCS and one LOW-CCS set with matched
mean query similarity ($|\Delta \simq|\leq0.02$). This directly tests H1:
whether coherence improves faithfulness after local relevance is held fixed.

\paragraph{Test 2: scaling the refinement effect.}
We rerun the SQuAD/Mistral headline cell at $n=500$ over five seeds and three
conditions: dense baseline, HCPC-v1 refinement, and HCPC-v2 gated refinement.
This tests whether the original pilot-sized refinement paradox remains large
and stable at scale.

\paragraph{Test 3: changing the faithfulness observer.}
The same 7,500 generations are scored with DeBERTa-v3 NLI, a second NLI model,
and a local RAGAS-style judge. This keeps the model outputs fixed and varies
only the observer, isolating metric dependence.

\paragraph{Test 4: transferring policy thresholds.}
We tune the HCPC threshold $\tau$ on each of five datasets and evaluate it on
all five. A robust policy should not require large in-domain tuning advantage.

\paragraph{Test 5: separating coherence from generic noise.}
We compare random off-topic noise with same-topic but answer-uninformative
noise. If coherence is only another name for retrieval quality, the two noise
types should have similar faithfulness slopes. If coherence carries diagnostic
signal beyond generic degradation, same-topic uninformative noise should be
less harmful.

\paragraph{Test 6: comparing deployment policies.}
We evaluate HCPC-v2, CRAG, RAPTOR-2L, and Self-RAG on SQuAD and HotpotQA at
$n=200$ per dataset/condition. The verified full Self-RAG artifact contains
1,600 per-query evaluations, 8 summary rows, and 0 error rows. We report
faithfulness, hallucination rate, mean latency, p99 latency, and indexing
cost. A separate RAPTOR audit reports p50/p99 latency, dense-index cost,
tree-build cost, and index size across SQuAD, PubMedQA, and HotpotQA.

\subsection{ControlledRAG as a Reusable Protocol}

\method{} is intended as an audit protocol for RAG claims, not as a one-off
set of experiments for this paper. The unit of analysis is a claim made by a
RAG system paper: that a retrieval change improves support, that a threshold
generalizes, that a metric identifies an effect size, or that a method is
deployment-preferable. Each protocol step asks what else would need to be true
for that claim to be identified.

\begin{table}[H]
\centering
\footnotesize
\caption{\method{} audit controls. Each control can be reused with other
datasets, retrievers, generators, and faithfulness metrics.}
\label{tab:controlledrag_protocol}
\begin{tabularx}{\linewidth}{p{0.20\linewidth}p{0.24\linewidth}Xp{0.20\linewidth}}
\toprule
Control & Claim it tests & Failure mode exposed & Primary artifact/table \\
\midrule
Similarity-matched intervention & Set coherence improves faithfulness after
local relevance is fixed & A plausible set statistic can lose its apparent
effect once local query--passage similarity is controlled & Table~\ref{tab:causal};
\texttt{paired\_wilcoxon.csv} \\
Multi-seed scale correction & A pilot effect remains large under repeated
sampling & Small effects can be seed-sensitive even when the direction looks
stable in a small headline cell & Table~\ref{tab:scaled};
\texttt{headline\_table.csv} \\
Fixed-generation multi-metric scoring & A single automatic judge identifies the
intervention effect size & The same generations can imply different effect
magnitudes under different faithfulness observers & Table~\ref{tab:metrics};
\texttt{table1} multimetric CSV \\
Threshold-transfer matrix & One tuned threshold generalizes across datasets &
In-domain tuning can hide poor off-diagonal transfer & Table~\ref{tab:tauflags};
\texttt{tau} transfer matrix CSV \\
Coherent-vs-random noise separation & Coherence is just generic retrieval
quality under another name & Same-topic uninformative context and random
off-topic context can have different faithfulness slopes & Table~\ref{tab:noise};
\texttt{slope\_response.csv} \\
Cost-aware Pareto audit & One method is deployment-preferable & A method can
win one dataset or metric while losing on p99 latency, hallucination, or
offline indexing cost & Figure~\ref{fig:pareto_p99};
Table~\ref{tab:h2h} \\
\bottomrule
\end{tabularx}
\end{table}

The protocol is deliberately modular. A future RAG paper can replace the
datasets, generators, retrievers, or judges while keeping the same audit
questions: match the confounder, scale the headline cell, rescore fixed
generations, transfer thresholds, separate noise types, and report the cost
surface.
