\section{Experimental Setup}
\label{sec:setup}

We hold the entire pipeline fixed across experiments and vary only one
factor at a time. This is intentional: the paradox is a property of the
\emph{interaction} between the refinement step and the rest of the
pipeline, not of any specific configuration.

\paragraph{Datasets.}
We evaluate on five public QA benchmarks: SQuAD~\citep{rajpurkar2016squad}
(general-domain extractive), PubMedQA~\citep{jin2019pubmedqa}
(biomedical), HotpotQA~\citep{yang2018hotpotqa} (multi-hop),
NaturalQuestions~\citep{kwiatkowski2019nq}, and TriviaQA. Each provides
a query, a ground-truth answer span, and a passage corpus. We
randomly sample $30$ queries per (dataset, condition) cell unless
otherwise noted, fixed across conditions to remove query-difficulty
confound.

\paragraph{Generators.}
Five model families covering $7$B--$120$B parameter scale: Mistral-$7$B
and Llama-$3$-$8$B (default open-weight tier); Qwen-$2.5$-$7$B
(alternative family); Llama-$3.3$-$70$B (frontier comparison); and
GPT-OSS-$120$B (largest open-weight model in Groq's catalog). All
served either via Ollama (local) or via the Groq free-tier API.

\paragraph{Retrieval stack.}
Embeddings come from \texttt{sentence-transformers/all-MiniLM-L$6$-v$2$};
indexing uses ChromaDB; refinement uses cross-encoder/ms-marco-MiniLM-L-$6$-v$2$.
Default chunk size is $1024$ characters with $10\%$ overlap; default top-$k$
is $3$ unless explicitly varied (\S\ref{sec:robustness}).

\paragraph{Conditions.}
Three retrieval configurations evaluated head-to-head per query:
\textbf{baseline} (top-$k$ dense retrieval, no refinement);
\textbf{HCPC-v$1$} (refine all retrieved passages via the cross-encoder
pipeline --- the failure mode we attack); and
\textbf{HCPC-v$2$} (coherence-gated refinement; described in
\S\ref{sec:method}).

\paragraph{Metrics.}
Per-query \textbf{faithfulness} is the cross-encoder NLI entailment
probability between the generated answer and the retrieved context;
\textbf{hallucination} is the indicator that faithfulness $<0.5$;
\textbf{retrieval similarity} is the mean cosine similarity between
the query and each retained chunk; \textbf{Context Coherence Score
(CCS)} is the mean minus standard deviation of off-diagonal
pairwise cosine similarities between retained chunk embeddings.

\paragraph{Statistical protocol.}
Headline numbers are reported with $3$-seed cross-seed standard
deviations (\S\ref{sec:robustness}). Cross-condition comparisons use
paired Wilcoxon signed-rank tests at $\alpha=0.05$. ANOVA and
Cohen's $d$ are reported in the appendix.

\paragraph{Reproducibility.}
Code, all per-query CSVs, the compiled benchmark, and a Docker image
are released; the benchmark also lives at
\href{https://huggingface.co/datasets/saketmgnt/context-coherence-bench}
{\texttt{saketmgnt/context-coherence-bench}} and as a permanent DOI
\href{https://doi.org/10.5281/zenodo.19757291}{\texttt{10.5281/zenodo.19757291}}.
A standalone $\texttt{pip install context-coherence}$ package ships
the metric and the deployment policy.
