\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage{natbib}
\usepackage[section]{placeins}

\title{\Large\bfseries Supplementary Material:\\
When Retrieval Quality Decouples from Faithfulness:\\
A Pre-Registered Audit of RAG Evaluation}
\author{Anonymous Authors\\\textit{Submission under double-blind review}}
\date{}

\begin{document}
\maketitle

\tableofcontents

\section{Pre-Registration}
\label{supp:prereg}

The matched-context, scaled, metric-fragility, threshold-transfer,
noise, cost-aware-baseline, and human-calibration audits were
pre-registered on 2026-04-26 before any results were inspected. (The
human-calibration design---stratified $n=100$, two annotators, Cohen's
$\kappa$---was specified as the optional human-annotation arm of the
metric-fragility pre-registration, and was executed at $n=99$ within
two days of the freeze.) The pre-registration
fixed: the matched-pair construction (similarity gap $\leq 0.02$, CCS
gap $\geq 0.20$); the paired Wilcoxon test direction (HIGH$>$LOW); the
$10{,}000$-resample bootstrap for the CI on the mean paired difference;
the Wilson $95\%$ CI on hallucination rates; the seed list
$\{41, 42, 43, 44, 45\}$; the per-seed paired Wilcoxon as the primary
significance test in the scaled audit; the three faithfulness scorers
(DeBERTa proxy, \texttt{roberta-large-mnli} proxy, RAGAS-style) plus the
fixed-context rescore protocol; the $\tau$-sweep range; and the
$200$-query head-to-head sample size. The four follow-up arms were
added after the pre-registered runs as post-hoc robustness checks:
answer-span/control diagnostic, stronger-retriever sanity check,
Qwen$2.5$ second-generator replication, and long-form stress test. We disclose which experiments are
pre-registered (\texttt{audit\_log.md} sections under
``Pre-Registered'') versus follow-ups (\texttt{audit\_log.md} sections
under ``Follow-Up'') in the file
\texttt{audit\_log.md} shipped at the top level of the artifact.

\section{Scaled Audit: Per-Seed Detail}
\label{supp:scaled}

The scaled audit on SQuAD/Mistral runs $n=2{,}500$ examples per
condition for each of five seeds; pooled means appear in the main
paper. Table~\ref{tab:supp:scaled_per_seed} reports per-seed paired
Wilcoxon $p$-values, mean paired differences, and Cohen's $d_z$ for
each contrast.

\begin{table}[!htb]
\centering
\caption{Per-seed paired contrasts on SQuAD/Mistral, $n=500$ pairs per
seed. Significance is paired Wilcoxon at $\alpha=0.05$.}
\label{tab:supp:scaled_per_seed}
\footnotesize
\begin{tabular}{lcccc}
\toprule
Seed & Contrast & Mean diff & Wilcoxon $p$ & Cohen's $d_z$ \\
\midrule
41 & baseline$-$v$1$ & $+0.0107$ & $0.083$ & $+0.072$ \\
42 & baseline$-$v$1$ & $+0.0063$ & $0.234$ & $+0.043$ \\
43 & baseline$-$v$1$ & $+0.0078$ & $0.174$ & $+0.055$ \\
44 & baseline$-$v$1$ & $+0.0204$ & $\mathbf{0.008}$ & $+0.134$ \\
45 & baseline$-$v$1$ & $+0.0073$ & $0.203$ & $+0.048$ \\
\midrule
41 & v$2$$-$v$1$ & $+0.0132$ & $\mathbf{0.025}$ & $+0.088$ \\
42 & v$2$$-$v$1$ & $+0.0030$ & $0.504$ & $+0.021$ \\
43 & v$2$$-$v$1$ & $+0.0059$ & $0.215$ & $+0.042$ \\
44 & v$2$$-$v$1$ & $+0.0191$ & $\mathbf{0.008}$ & $+0.126$ \\
45 & v$2$$-$v$1$ & $+0.0125$ & $\mathbf{0.036}$ & $+0.081$ \\
\bottomrule
\end{tabular}
\end{table}

Bootstrap CIs on the pooled mean paired difference for
baseline$-$v$1$ include zero, and only $1$ of $5$ seeds reaches
significance for that contrast. The v$2$$-$v$1$ contrast reaches
significance in $3$ of $5$ seeds, but the magnitude is small (per-seed
mean diff $0.003$--$0.019$). Seed-level uncertainty is treated as a
ControlledRAG reporting axis, not a result to average away.

\section{Answer-Span/Control Diagnostic Details}
\label{supp:span}

The answer-span detector lowercases text, strips punctuation,
normalises whitespace, and checks whether any normalised gold-answer
alias appears in the retrieved context for the audited RAG
setting~\citep{lewis2020rag}. It uses no external model. The resulting
control rows and model summaries are written under the Fix 12 data and
results directories.

\begin{table}[!htb]
\centering
\caption{Matched-context span and outcome summary.}
\footnotesize
\begin{tabular}{lrrrrr}
\toprule
Group & $n$ & span present & faithfulness & hallucination & CCS \\
\midrule
all          & $400$ & $0.220$ & $0.637$ & $0.128$ & $0.404$ \\
HIGH-CCS     & $200$ & $0.170$ & $0.636$ & $0.165$ & $0.670$ \\
LOW-CCS      & $200$ & $0.270$ & $0.639$ & $0.090$ & $0.137$ \\
span absent  & $312$ & $0.000$ & $0.628$ & $0.119$ & $0.408$ \\
span present & $\phantom{1}88$ & $1.000$ & $0.670$ & $0.159$ & $0.390$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!htb]
\centering
\caption{Full span/control model comparison. CIs are bootstrap $95\%$
intervals from the script output.}
\footnotesize
\begin{tabular}{lrrrr}
\toprule
Predictors & $R^2$ & $R^2$ CI & AUROC & AUROC CI \\
\midrule
answer span                   & $0.019$ & $[0.001, 0.062]$ & $0.531$ & $[0.502, 0.599]$ \\
mean similarity               & $0.002$ & $[0.000, 0.023]$ & $0.586$ & $[0.495, 0.677]$ \\
CCS                           & $0.003$ & $[0.000, 0.022]$ & $0.639$ & $[0.564, 0.712]$ \\
span $+$ similarity           & $0.024$ & $[0.004, 0.071]$ & $0.590$ & $[0.518, 0.687]$ \\
span $+$ CCS                  & $0.022$ & $[0.003, 0.068]$ & $0.631$ & $[0.572, 0.716]$ \\
span $+$ similarity $+$ CCS   & $0.025$ & $[0.006, 0.075]$ & $0.643$ & $[0.589, 0.734]$ \\
\bottomrule
\end{tabular}
\end{table}

The headline reading is reported in the main paper: span presence is
the strongest single predictor of continuous faithfulness, CCS is the
strongest single predictor of hallucination AUROC, and the joint
feature set has the highest score on both. The $R^2$ values are uniformly small in
absolute terms; the relative ranking is the load-bearing signal.
Span presence is a weak lexical proxy, not a gold support variable:
retrieved text can contain the answer string while still failing to
support the generated answer. The mean-score and thresholded
hallucination views also inspect different parts of the score
distribution.

\begin{table}[!htb]
\centering
\caption{Matched HIGH/LOW CCS faithfulness distribution.}
\footnotesize
\begin{tabular}{lrrrrrrr}
\toprule
Group & $n$ & mean & p10 & p25 & p50 & p75 & p90 \\
\midrule
HIGH-CCS & $200$ & $0.636$ & $0.476$ & $0.553$ & $0.638$ & $0.713$ & $0.795$ \\
LOW-CCS  & $200$ & $0.639$ & $0.512$ & $0.569$ & $0.648$ & $0.715$ & $0.769$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!htb]
\centering
\caption{Matched HIGH/LOW CCS hallucination rates as the score
threshold varies.}
\footnotesize
\begin{tabular}{lrrrrr}
\toprule
Group & $\theta=0.3$ & $\theta=0.4$ & $\theta=0.5$ & $\theta=0.6$ & $\theta=0.7$ \\
\midrule
HIGH-CCS & $0.0\%$ & $3.0\%$ & $16.5\%$ & $38.5\%$ & $71.0\%$ \\
LOW-CCS  & $1.0\%$ & $3.0\%$ & $\phantom{1}9.0\%$ & $38.5\%$ & $70.0\%$ \\
\bottomrule
\end{tabular}
\end{table}

These tables explain why the matched mean contrast can be null while a
specific thresholded hallucination contrast differs: the distributions
are close in the middle, and the observed HIGH/LOW difference is
threshold-sensitive rather than a clean directional CCS effect.

\section{Stronger Retriever Sanity Check}
\label{supp:retriever}

The retriever sanity check has two parts. First, the matched contexts
from the main-paper audit are re-encoded with MiniLM, BGE-large, and
E5-large; this is a fixed-context diagnostic, not a fresh generation
run. Second, the completed multi-retriever generation artifact is
copied into \texttt{data/revision/fix\_13/} and summarised. The fixed
generations are unchanged across embedders, so the matched
HIGH$-$LOW faithfulness contrast inherits the matched-context null
from the main paper and is not re-tested here.

\begin{table}[!htb]
\centering
\caption{Fixed-context re-encoding of the matched HIGH/LOW CCS pairs.
The HIGH$>$LOW CCS ordering is preserved across all three embedders.}
\footnotesize
\begin{tabular}{lrrrrr}
\toprule
Embedder & pairs & CCS high & CCS low & high$-$low & sim gap \\
\midrule
MiniLM    & $200$ & $0.670$ & $0.137$ & $0.533$ & $0.006$ \\
BGE-large & $200$ & $0.739$ & $0.491$ & $0.249$ & $0.032$ \\
E5-large  & $200$ & $0.830$ & $0.702$ & $0.127$ & $0.015$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!htb]
\centering
\caption{Paradox magnitude by embedding model from the completed
multi-retriever generation artifact ($n=30$ per cell, original pilot
sampling protocol re-run under different embedders). The SQuAD column
shows that the paradox direction depends on retriever; PubMedQA
preserves a positive paradox across all four embedders.
\textit{These rows are not directly comparable to the $n=2{,}500$
scaled-audit number in main paper \S6 (baseline$-$v$1$ $= 0.011$):
that table uses the audit-construction sampling and a single MiniLM
retriever, whereas this table uses the small-$n$ pilot sampling and
varies the embedder. Cross-table comparison should be same-row across
embedders within this table, not this table vs.\ the main-paper
scaled-audit table.}}
\footnotesize
\begin{tabular}{llrr}
\toprule
Dataset & embedder & baseline$-$v$1$ & v$2$$-$v$1$ \\
\midrule
SQuAD    & MiniLM    & $+0.075$ & $+0.099$ \\
SQuAD    & BGE-large & $+0.032$ & $+0.025$ \\
SQuAD    & E5-large  & $-0.011$ & $+0.008$ \\
SQuAD    & GTE-large & $-0.001$ & $-0.006$ \\
PubMedQA & MiniLM    & $+0.024$ & $+0.019$ \\
PubMedQA & BGE-large & $+0.033$ & $+0.028$ \\
PubMedQA & E5-large  & $+0.053$ & $+0.039$ \\
PubMedQA & GTE-large & $+0.006$ & $+0.002$ \\
\bottomrule
\end{tabular}
\end{table}

The safe interpretation is retriever sensitivity, not universal
robustness. Stronger embeddings change the retrieval surface, sometimes
in a direction that flips the SQuAD paradox sign. ControlledRAG asks
for retriever identity to be reported as a routine audit axis;
the supplement here is the operational example.

\section{Second-Generator Metric-Fragility Replication}
\label{supp:qwen}

The second-generator audit samples $n=100$ fixed contexts per condition
from the matched-context construction and generates answers with
Qwen$2.5$ via local Ollama. Those fixed generations are scored with
the DeBERTa proxy, the \texttt{roberta-large-mnli} proxy, and a local
RAGAS-style judge (Mistral via Ollama).

\begin{table}[!htb]
\centering
\caption{Qwen$2.5$ fixed-generation metric means and raw score
contrasts.}
\footnotesize
\begin{tabular}{lrrrrr}
\toprule
Metric & baseline & HCPC-v$1$ & HCPC-v$2$ & baseline$-$v$1$ & v$2$$-$v$1$ \\
\midrule
DeBERTa             & $0.645$ & $0.631$ & $0.644$ & $+0.015$ & $+0.014$ \\
roberta-large-MNLI  & $0.362$ & $0.316$ & $0.368$ & $+0.046$ & $+0.051$ \\
RAGAS-style         & $0.806$ & $0.672$ & $0.814$ & $+0.134$ & $+0.142$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!htb]
\centering
\caption{Metric correlations on the Qwen$2.5$ fixed-generation rows
($n=300$).}
\footnotesize
\begin{tabular}{lrr}
\toprule
Metric pair & Pearson $r$ & Spearman $\rho$ \\
\midrule
DeBERTa vs.\ roberta-large-MNLI       & $0.380$ & $0.429$ \\
DeBERTa vs.\ RAGAS-style              & $0.280$ & $0.318$ \\
roberta-large-MNLI vs.\ RAGAS-style   & $0.721$ & $0.668$ \\
\bottomrule
\end{tabular}
\end{table}

The scorer-disagreement ordering is preserved under this compact
second-generator replication. The claim is limited to this audited
Qwen$2.5$ cell; ControlledRAG explicitly asks for a broader generator
panel before any universal-generator claim is made.

\section{Scorer Implementation and Standardized Contrasts}
\label{supp:scorer_standardized}

The frozen artifact uses two legacy NLI-style zero-shot label proxies
and one context-conditioned RAGAS-style judge. The DeBERTa proxy
(\nolinkurl{cross-encoder/nli-deberta-v3-base}) and RoBERTa proxy
(\nolinkurl{roberta-large-mnli}) score each generated answer sentence
as the sequence to be classified against the labels
entailment/neutral/contradiction with
\texttt{hypothesis\_template="\{\}"}. The retrieved context is retained
in the per-row data and used by the RAGAS-style judge, but it is not
consumed by those two legacy zero-shot proxy calls. This disclosure is
part of the scorer audit: model identity, input formatting, and prompt
format are not implementation trivia.

Because the three scorers are on different native scales, the main
paper reports raw contrasts as score-scale-dependent contrasts and
adds a standardized Mistral analysis. Table~\ref{tab:supp:standardized}
z-scores each scorer over the same $7{,}500$ fixed-generation rows and
also reports percentile-rank-normalized contrasts. CIs are paired
bootstrap intervals over matched query/seed keys; $2{,}499$ paired keys
are available after one duplicated key is collapsed by the pivot.

\begin{table}[!htb]
\centering
\caption{Standardized fixed-generation scorer-fragility contrasts on
Mistral rows. Positive values mean the left condition scores higher
than HCPC-v$1$.}
\label{tab:supp:standardized}
\scriptsize
\begin{tabular}{llrrr}
\toprule
Scorer & Contrast & raw diff & z diff (CI) & rank diff (CI) \\
\midrule
DeBERTa proxy & baseline$-$v$1$ & $0.010$ & $0.071\,[0.032,0.110]$ & $0.022\,[0.010,0.033]$ \\
DeBERTa proxy & v$2$$-$v$1$ & $0.011$ & $0.073\,[0.034,0.112]$ & $0.022\,[0.011,0.034]$ \\
roberta-large-MNLI proxy & baseline$-$v$1$ & $0.032$ & $0.231\,[0.191,0.271]$ & $0.062\,[0.051,0.074]$ \\
roberta-large-MNLI proxy & v$2$$-$v$1$ & $0.032$ & $0.236\,[0.199,0.275]$ & $0.063\,[0.052,0.075]$ \\
RAGAS-style judge & baseline$-$v$1$ & $0.139$ & $0.337\,[0.298,0.375]$ & $0.085\,[0.075,0.095]$ \\
RAGAS-style judge & v$2$$-$v$1$ & $0.138$ & $0.332\,[0.294,0.370]$ & $0.083\,[0.073,0.092]$ \\
\bottomrule
\end{tabular}
\end{table}

The standardized table preserves the scorer-fragility conclusion
without treating raw scores on different scales as directly comparable
causal effects.

\section{Long-Form/Synthesis Stress Test}
\label{supp:longform}

The long-form stress test packages an existing QASPER/MS-MARCO run
with $40$ distinct long-form questions. It is included only as an
exploratory scope probe, not as broad generalization evidence for the
short-answer audit.

\begin{table}[!htb]
\centering
\caption{Long-form stress-test condition means.}
\footnotesize
\begin{tabular}{llrrrr}
\toprule
Dataset & condition & span faith & claim faith & unsupported & long halluc.\ \\
\midrule
MS-MARCO & baseline   & $0.645$ & $0.646$ & $0.224$ & $0.400$ \\
MS-MARCO & HCPC-v$1$  & $0.678$ & $0.679$ & $0.075$ & $0.050$ \\
MS-MARCO & HCPC-v$2$  & $0.674$ & $0.674$ & $0.158$ & $0.250$ \\
QASPER   & baseline   & $0.649$ & $0.649$ & $0.158$ & $0.300$ \\
QASPER   & HCPC-v$1$  & $0.647$ & $0.647$ & $0.117$ & $0.200$ \\
QASPER   & HCPC-v$2$  & $0.648$ & $0.651$ & $0.129$ & $0.250$ \\
\bottomrule
\end{tabular}
\end{table}

The short-answer paradox does not generalise cleanly here: MS-MARCO has
a negative span-faithfulness paradox drop ($-0.033$), and QASPER is
near zero ($+0.002$). The long-form result is reported as a scope
limit rather than as broad long-form evidence for ControlledRAG.

\section{Self-RAG: Harness-Mismatched Appendix Baseline}
\label{supp:selfrag}

We include Self-RAG~\citep{asai2024selfrag} only as an appendix
baseline because the available Self-RAG harness mismatches the audit's
fixed-context protocol along three dimensions: (i) Self-RAG's
generator emits its own retrieval and reflection tokens, so the
``fixed retrieved context'' assumption used by the matched-context and
metric-fragility audits is not directly satisfied; (ii) Self-RAG's
recommended generator is a fine-tuned Llama-$2$-$7$B variant that does
not coincide with the audited Mistral-$7$B / Qwen$2.5$-$7$B family
and that was OOM-killed on the available T$4$ instance for the
PubMedQA cell, so the partial SQuAD numbers we have are not directly
comparable to the cost-aware head-to-head; and (iii) the
recommended Self-RAG threshold differs from the threshold-transfer
sweep used elsewhere. We report this disclosure here so that future
ControlledRAG-style audits can budget for a matched Self-RAG harness;
we explicitly do not generalize from these mismatched runs to Self-RAG
overall.

\section{Coherence vs.\ Generic Retrieval Noise}
\label{supp:noise}

The noise audit uses the matched-context construction as a backbone
and replaces $1/3$, $2/3$, or $3/3$ of the retrieved chunks with one
of two noise types: random off-topic passages and coherence-preserving
uninformative passages drawn from the same topic distribution that do
not contain the answer.

\begin{table}[!htb]
\centering
\caption{Faithfulness and similarity slopes per noise rate.}
\footnotesize
\begin{tabular}{lrr}
\toprule
Condition & faith slope (per noise rate) & sim slope (per noise rate) \\
\midrule
random noise                     & $-0.069$ & $-0.481$ \\
coherent uninformative noise     & $-0.043$ & $-0.113$ \\
HCPC-v$1$ refinement (reference) & ---      & ---      \\
\bottomrule
\end{tabular}
\end{table}

The two noise types produce distinguishable faithfulness slopes at
matched substitution rate; we do not claim the gap is large in absolute
terms. The right reading is that coherence and similarity are
correlated but not identical.

\section{Threshold Transfer: Full Sweep}
\label{supp:threshold}

The main paper reports a per-tune-dataset summary
(diagonal recovery, off-diagonal mean, gap) at each tune dataset's
selected $\tau^\ast$. The full $5 \times 5$ per-cell transfer matrix at
each tune dataset's $\tau^\ast$ is reproduced below; the per-$\tau$
sweep over $\{0.3, 0.4, 0.5, 0.6, 0.7\}$ for each dataset is in the
artifact's \texttt{results/revision/fix\_04/} directory.
Recovery is the dimensionless normalized contrast
\[
R_\tau(d)=
\frac{\mathrm{faith}_{\mathrm{gate},\tau}(d) -
      \mathrm{faith}_{\mathrm{HCPC}\text{-}v1}(d)}
     {\mathrm{faith}_{\mathrm{baseline}}(d) -
      \mathrm{faith}_{\mathrm{HCPC}\text{-}v1}(d)}.
\]
It is useful for asking whether a tuned threshold transports, but it is
not a direct utility measure: when the denominator is small or negative,
the sign can be counterintuitive and values can exceed $1$.

\begin{table}[!htb]
\centering
\caption{Full $5 \times 5$ threshold-transfer matrix. Each row gives
the normalized recovery ratio of the $\tau$ tuned on the row dataset,
evaluated on each of the five column datasets. Positive values are not
always ``better'' when the baseline$-$v$1$ denominator is negative.
Rows can be byte-identical across the eval columns when two tune
datasets pick the same $\tau^\ast$: tune-SQuAD and tune-HotpotQA both
pick $\tau^\ast = 0.3$, and tune-PubMedQA and tune-TriviaQA both pick
$\tau^\ast = 0.4$, so the off-diagonal columns coincide. This is a
faithful consequence of the experiment (off-diagonal recovery only
depends on the eval dataset and the selected $\tau$, not on which tune
dataset selected it), not a pipeline bug.}
\label{tab:supp:threshold_transfer_full}
\footnotesize
\begin{tabular}{lcrrrrr}
\toprule
$\tau$ tuned on & $\tau^\ast$ & SQuAD & PubMedQA & HotpotQA & NaturalQs & TriviaQA \\
\midrule
PubMedQA  & $0.4$ & $+0.489$ & $+1.452$ & $-0.291$ & $+1.140$ & $+0.464$ \\
NaturalQs & $0.5$ & $+0.327$ & $+0.664$ & $-0.471$ & $+1.401$ & $+0.323$ \\
SQuAD     & $0.3$ & $+0.823$ & $-1.714$ & $-0.112$ & $+0.865$ & $+0.398$ \\
TriviaQA  & $0.4$ & $+0.489$ & $+1.452$ & $-0.291$ & $+1.140$ & $+0.464$ \\
HotpotQA  & $0.3$ & $+0.823$ & $-1.714$ & $-0.112$ & $+0.865$ & $+0.398$ \\
\bottomrule
\end{tabular}
\end{table}

Per-$\tau$ recovery curves and confidence diagnostics for the
$\tau$-tuning step are included as machine-readable CSVs in the same
results directory. The transfer matrix itself does not include
bootstrap CIs, so we use it as a descriptive transport diagnostic.

\section{Human Calibration Protocol}
\label{supp:human}

The human-calibration panel asks two raters (one expert, one
general-domain) to label $n=99$ matched-cell items with a binary
faithful/hallucinated decision and a $0$--$1$ confidence score. The
panel is designed for calibration of automated scorers against a
reference human label set on the audited cell, not for a settled scorer
ranking.

\paragraph{Protocol.}
We use rater~$1$'s labels as the reference human label set; rater~$2$
independently re-labeled the same $99$ items as a reliability check.
Cohen's $\kappa = 0.774$ between the two label sets is reported as
inter-rater reliability rather than as a disagreement-resolution
process. We do not collapse the two raters into a third
``resolved'' label by majority or by discussion; with two raters
either procedure would be ill-defined. Spearman correlations against
the rater-$1$ label set are
$\rho_{\textrm{DeBERTa}} = 0.140$,
$\rho_{\textrm{second NLI}} = 0.380$,
$\rho_{\textrm{RAGAS-style}} = 0.441$. Bootstrap $95\%$ CIs on the
three correlations overlap, so the metric ranking is suggestive rather
than settled.

\section{Source Trace and Artifact Manifest}
\label{supp:trace}

Every numerical claim in the main paper traces to a result file under
\texttt{results/revision/} with a corresponding pre-registered script
under \texttt{experiments/}. Table~\ref{tab:supp:trace} summarises the
trace; the full mapping is shipped as \texttt{SOURCE\_TRACE.md} in the
artifact and reproduced here for the reviewer's convenience.

\begin{table}[!htb]
\centering
\caption{Audit trace from claim to source file.}
\label{tab:supp:trace}
\footnotesize
\begin{tabular}{ll}
\toprule
Audit & Source results files \\
\midrule
Matched-context (\S\ref{supp:span} backbone) &
  \texttt{results/revision/fix\_01/} \\
Scaled audit (\S\ref{supp:scaled}) &
  \texttt{results/revision/fix\_02/} \\
Metric fragility (Mistral) &
  \texttt{results/revision/fix\_03/}, \texttt{standardized\_scorer\_fragility.csv} \\
Threshold transfer (\S\ref{supp:threshold}) &
  \texttt{results/revision/fix\_04/} \\
Noise (\S\ref{supp:noise}) &
  \texttt{results/revision/fix\_05/} \\
Cost-aware head-to-head &
  \texttt{results/revision/fix\_06/}, \texttt{h2h\_summary\_with\_ci.csv}, \texttt{fix\_11/} \\
Span/control (\S\ref{supp:span}) &
  \texttt{results/revision/fix\_12/} \\
Stronger retrievers (\S\ref{supp:retriever}) &
  \texttt{results/revision/fix\_13/}, \texttt{results/multi\_retriever/} \\
Second generator (\S\ref{supp:qwen}) &
  \texttt{results/revision/fix\_14/} \\
Long-form stress (\S\ref{supp:longform}) &
  \texttt{results/revision/fix\_15/} \\
Human calibration (\S\ref{supp:human}) &
  \texttt{results/human\_*/} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Reproducibility.} An anonymized artifact package containing
code, per-query CSVs, matched context pairs, source tables, scripts, and
documentation is provided for review. Public package, benchmark, and
permanent archive links are omitted during double-blind review and will
be released upon acceptance.

\section{Resume and Compute Notes}
\label{supp:kaggle}

The local scripts are resumable through their CSV outputs. The only
script likely to benefit from a free-tier GPU notebook is the
second-generator replication if local Ollama is slow: run the same
command on a free GPU-backed notebook after pulling Qwen$2.5$ and
Mistral into Ollama, then download \texttt{data/revision/fix\_14/} and
\texttt{results/revision/fix\_14/}. None of the audit experiments
require paid APIs; all scorer rescoring is local DeBERTa,
\texttt{roberta-large-mnli}, and a Mistral-via-Ollama RAGAS-style
judge.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
