% Fix 6 - baseline head-to-head.

\subsection{Head-to-Head Baselines}
\label{sec:fix06_baselines}

We compare HCPC-v2 with CRAG and a two-level RAPTOR implementation in the
same generation/scoring harness on \textsc{SQuAD} and HotpotQA. The table
reports faithfulness, hallucination rate, p99 latency, per-1k-query wall
clock cost, and indexing cost. A faithfulness-vs-latency Pareto plot
summarizes the deployment trade-off.

The no-Self-RAG Kaggle T4$\times$2 run completed with 1,200 evaluations:
200 queries per dataset for each of HCPC-v2, CRAG, and RAPTOR-2L. The
downloaded package does not include Self-RAG smoke or full-run outputs, so
Self-RAG remains an optional follow-up rather than a populated row.

\begin{table}[t]
\centering
\small
\begin{tabular}{llrrrrr}
\toprule
Dataset & Method & $n$ & Faith. & Halluc. & Mean lat. & p99 lat. \\
 & & & & & (ms) & (ms) \\
\midrule
HotpotQA & CRAG & 200 & 0.643 & 0.105 & 1941 & 4255 \\
HotpotQA & HCPC-v2 & 200 & 0.633 & 0.125 & 2289 & 5545 \\
HotpotQA & RAPTOR-2L & 200 & 0.631 & 0.130 & 2414 & 5074 \\
\textsc{SQuAD} & CRAG & 200 & 0.698 & 0.120 & 1440 & 5111 \\
\textsc{SQuAD} & HCPC-v2 & 200 & 0.708 & 0.125 & 1720 & 5400 \\
\textsc{SQuAD} & RAPTOR-2L & 200 & 0.710 & 0.125 & 1714 & 4442 \\
\bottomrule
\end{tabular}
\caption{Fix 6 no-Self-RAG head-to-head comparison. HCPC-v2, CRAG, and
RAPTOR-2L are evaluated under the same generator and scoring harness.}
\label{tab:fix06_noselfrag}
\end{table}

The comparison does not support a blanket dominance claim for HCPC-v2. On
\textsc{SQuAD}, HCPC-v2 and RAPTOR-2L are effectively tied in faithfulness
and hallucination rate, with RAPTOR-2L ahead by only 0.0016 faithfulness
points. On HotpotQA, CRAG is strongest among the no-Self-RAG methods, with
higher faithfulness, lower hallucination rate, and lower mean latency than
HCPC-v2 or RAPTOR-2L. RAPTOR also requires a substantial offline tree-build
cost: 134--175 seconds in this run versus roughly 8 seconds for the dense base
index. The paper should therefore frame HCPC-v2 as competitive in this
head-to-head rather than uniformly superior.
