\section{The Refinement Paradox}
\label{sec:paradox}

\input{figures/headline_frontier}

We open with the phenomenon. Table~\ref{tab:hcpc_main} reports the
baseline / HCPC-v$1$ / HCPC-v$2$ comparison on SQuAD with Mistral-$7$B
at the canonical operating point ($k{=}3$, chunk = $1024$ chars,
$n{=}30$). Naive per-passage refinement (HCPC-v$1$) raises mean
retrieval similarity from $0.587$ to $0.601$ --- the metric most RAG
papers optimize --- yet drops faithfulness from $0.7995$ to $0.6407$
and introduces $10.0\%$ hallucination where there was none.

\begin{table}[!htb]
\centering
\caption{The refinement paradox at the canonical operating point.
Mistral-$7$B, SQuAD, $k=3$, chunk $= 1024$, $n=30$ per condition.
HCPC-v$1$ \emph{maintains or improves} retrieval similarity while
\emph{degrading} faithfulness and introducing hallucination.
HCPC-v$2$ recovers faithfulness while intervening on $16.7\%$ of queries.}
\label{tab:hcpc_main}
\footnotesize
\begin{tabular}{lcccc}
\toprule
Condition & faithfulness & hallucination & retrieval~sim. & refine~rate \\
\midrule
Baseline       & $0.7995$ & $0.000$ & $0.587$ & $0.000$ \\
HCPC-v$1$ (refine all) & $0.6407$ & $\mathbf{0.100}$ & $0.601$ & $1.000$ \\
HCPC-v$2$ (gated)      & $\mathbf{0.7874}$ & $0.000$ & $0.589$ & $0.167$ \\
\bottomrule
\end{tabular}
\end{table}

The paradox is not specific to one (dataset, generator) cell. Across
the multidataset evaluation it appears on $11/15$ (dataset, generator)
pairs at $\Delta_{\textrm{paradox}} > 0.01$, with magnitude correlated
with the structural properties of the dataset rather than with the
generator family (Appendix~the multidataset evaluation (longform appendix)).

\subsection{The mechanism: it is the \emph{set}, not the passages}
\label{sec:paradox:mechanism}

The standard explanation for hallucination in RAG --- ``the passages
are wrong'' --- predicts that lower retrieval similarity should
correlate with worse faithfulness. We find the opposite: the worst
faithfulness drops happen when refinement \emph{raises} per-passage
similarity. The signal that drops alongside faithfulness is set-level
coherence: HCPC-v$1$'s mean CCS is $0.45$ vs the baseline's $0.59$
(the longform-paper appendix). The refinement step is breaking
the connective tissue between passages, not the per-passage relevance.

\input{figures/qualitative_paradox}

A representative example (Figure~\ref{fig:qualitative_paradox}) is
illustrative. The query ``\emph{What city did Super Bowl 50 take place
in?}'' has ground truth ``Santa Clara''. The baseline retrieval
returns three coherent chunks, all about Super Bowl 50 in Santa
Clara, and the generator answers correctly with faithfulness $0.94$.
HCPC-v$1$ refines each chunk to its query-most-similar fragment;
the refined set still has high per-chunk similarity to the query, but
the connective phrasing tying ``Santa Clara'' to ``Super Bowl 50'' is
gone. The generator falls back on the more familiar regional name and
answers ``San Francisco Bay Area'' --- fluent, plausible, and wrong
(faithfulness drops to $0.48$). HCPC-v$2$, which preserves the top
two chunks unconditionally, returns the correct ``Santa Clara'' answer.

\subsection{Disentangling coherence from relevance}
\label{sec:paradox:disentangle}

\input{figures/disentanglement}

A natural objection: maybe coherence is just relevance under another
name. Figure~\ref{fig:disentanglement} answers it directly. We bucket
queries into similarity quartiles (lines) and, within each quartile,
bin by CCS (x-axis). Faithfulness varies meaningfully along the
coherence axis \emph{at fixed similarity quartiles}: the spread within
a single similarity quartile is comparable to the spread across the
full data. If coherence were a re-statement of relevance, the lines
would be flat. They are not.


\subsection{What HCPC-v$2$ actually does}
\label{sec:paradox:v2_intuition}

HCPC-v$2$ is deliberately simple. After dense retrieval, it inspects
the top-$k$ retrieved set, computes CCS, and applies a single decision
rule: if CCS is below a threshold $\tau$ (a coherent set worth
preserving), refinement is skipped. Otherwise, the lowest-similarity
chunks are sub-divided and reranked, with the top two chunks
\emph{always} preserved by rank. The full algorithm and the bare-CCS
ablation that isolates the gate from the protected-rank rule are in
\S\ref{sec:method}.

The headline result: with this single decision rule, HCPC-v$2$
recovers SQuAD faithfulness from $0.6407$ back to $0.7874$ --- $93\%$
of the way back to the no-refinement baseline --- and brings
hallucination from $10\%$ to $0\%$. It does so while intervening on
only $16.7\%$ of queries: on the other $83.3\%$, it lets the standard
pipeline run unmodified. \emph{Restraint, not algorithmic
sophistication, is the active ingredient.}
