\section{Robustness}
\label{sec:robustness}

The paradox could plausibly be an artifact of any of seven design
choices we held fixed. We test all seven. The full per-experiment
tables are in the longform-paper appendix; this section presents the
distilled summary.

\begin{table}[!htb]
\centering
\caption{Seven robustness checks. Each row reports the headline
$\Delta_{\textrm{paradox}} = \textrm{faith}_{\textrm{base}} -
\textrm{faith}_{\textrm{v1}}$ and the HCPC-v$2$ recovery
$\Delta_{\textrm{recovery}} = \textrm{faith}_{\textrm{v2}} -
\textrm{faith}_{\textrm{v1}}$. The check is whether the result holds
in the direction the coherence account predicts.}
\label{tab:robustness_summary}
\footnotesize
\begin{tabular}{lllc}
\toprule
Axis & Variation tested & Headline finding & Verdict \\
\midrule
Multi-seed              & $3$ random seeds                    & SQuAD: $0.069 \pm 0.004$ ($17\times$ signal/$\sigma$)               & $\checkmark$ stable \\
Hierarchical (RAPTOR)   & RAPTOR-$2$L vs HCPC-v$2$            & HCPC wins $2/3$ datasets, ties on hallucination                    & $\checkmark$ competitive \\
Prompt template         & strict / cot / concise / expert     & ANOVA $p=0.83$ (SQuAD); $p=0.29$ (PubMedQA)                          & $\checkmark$ no main effect \\
Sub-chunk size          & $\{128, 256, 512\}$ tokens          & paradox visible in $3/3$ cells per dataset                          & $\checkmark$ stable \\
Frontier scale          & $7$B $\to$ $70$B $\to$ $120$B      & SQuAD/$70$B exactly reproduces $7$B magnitude ($+0.100$)            & $\checkmark$ persists \\
Top-$k$                 & $k \in \{2, 3, 5, 10\}$             & paradox \emph{grows} with $k$ on SQuAD ($0.063 \to 0.124$)          & $\checkmark$ predicted direction \\
Reranker family         & MiniLM-L$6$/L$12$/BAAI-bge          & paradox in all $6$ cells; recovery in all $6$                       & $\checkmark$ reranker-agnostic \\
Quantization            & Q$4$/Q$5$/Q$8$                      & SQuAD paradox $\in [0.088, 0.095]$; PubMedQA $\approx 0$ at all     & $\checkmark$ quant-agnostic \\
Temperature             & $T \in \{0, 0.3, 0.7, 1.0\}$        & paradox \emph{amplifies} with $T$ on SQuAD ($0.065 \to 0.115$)      & $\checkmark$ predicted direction \\
Long-form generation    & QASPER + MS-MARCO                   & paradox does not generalise; HCPC still cuts unsupported-rate $-67\%$ & \dag{} scope-qualified \\
Generic retrieval noise & inject $0$/$1$/$K$ off-topic        & SQuAD noise slope $> $ paradox magnitude                            & \dag{} qualitatively distinct only \\
\bottomrule
\end{tabular}
\end{table}

We highlight the four non-obvious entries.

\paragraph{Top-$k$ amplification.}
\input{figures/topk_table}
The coherence account predicts that increasing $k$ adds relevance
breadth (more matches) but spreads the retrieved set across more
disparate sources, so coherence should fall and the paradox should
grow. The full top-$k$ table confirms exactly this on SQuAD: the
paradox more than doubles from $k{=}2$ ($+0.063$) to $k{=}5$
($+0.124$). The bare CCS gate's recovery (\texttt{gate rec.}) tracks
HCPC-v$2$'s recovery within $\pm 0.05$ at every $k$, supporting the
claim that the gate decision is the active ingredient.

\paragraph{Frontier-scale persistence.}
The reviewer concern that the paradox is a small-model artifact is
addressed by holding the retrieval stack fixed and swapping only the
generator (Llama-$3.3$-$70$B and GPT-OSS-$120$B via Groq). On SQuAD,
Llama-$3.3$-$70$B exactly reproduces the $7$B-Mistral paradox magnitude
($0.100$ vs $0.100$), and GPT-OSS-$120$B still shows a positive
$+0.030$ paradox. PubMedQA paradox magnitudes shrink at scale
($0.005$/$0.013$ at $70$B/$120$B vs $0.025$ at $7$B), which the
coherence account predicts: stronger parametric priors are a more
plausible substitute for fragmented biomedical evidence than for
short-span SQuAD answers. The full table is in
the longform-paper appendix.

\paragraph{Temperature amplification.}
A novel finding from the temperature sweep:
$\textrm{faith}_{\textrm{base}} - \textrm{faith}_{\textrm{v1}}$ on
SQuAD grows monotonically with $T$, from $+0.065$ at $T=0$ to
$+0.115$ at $T=1.0$. The coherence interpretation is that higher $T$
samples more aggressively from the parametric prior, and a fragmented
context (HCPC-v$1$) gives the prior more room to operate. PubMedQA
paradox magnitudes stay flat across $T$ (consistent with the
domain-mismatch ceiling story). HCPC-v$2$ recovery is robust across
all $T$ and is largest at $T=0.7$ on SQuAD ($+0.130$).

\paragraph{Cross-validation: model self-confidence tracks CCS.}
We ask the same generator to rate its confidence ($0$--$100$) in the
answer it just produced, given the retrieved context. Pooling SQuAD +
PubMedQA ($n=60$), self-confidence is significantly correlated with
CCS: Pearson $r=0.36$ ($p=0.005$), Spearman $\rho=0.48$ ($p=0.0001$).
The model's internal coherence-tracking signal aligns with the
external diagnostic. Per-dataset correlations are weaker ($n=30$
each), so the result is to be read as a pooled cross-validation rather
than a strong per-dataset claim. Details in
the longform-paper appendix.

clusters are tight under baseline, scatter under HCPC-v$1$, and
recover under HCPC-v$2$.

\paragraph{Honest qualifications.}
Two robustness checks do \emph{not} land where the strongest version of
our claim would predict, and we report both transparently. First, the
paradox does not generalise to long-form generation on QASPER or
MS-MARCO; HCPC-v$2$ instead reduces the unsupported-claim rate
($-67\%$ on MS-MARCO). Second, on SQuAD the noise-injection slope
exceeds the paradox magnitude, so we cannot claim the paradox is
\emph{magnitudinally larger} than generic retrieval noise --- only
that it is sign-distinct from noise on the multi-hop and biomedical
regimes. Both qualifications are discussed in
\S\ref{sec:discussion} and elaborated in
the longform-paper appendix.
