Retrieval-augmented generation (RAG) is increasingly deployed in
high-stakes settings, yet evaluation practices vary substantially
across the literature: small samples, single faithfulness scorers,
and unreported hyperparameter tuning are common. We present a
pre-registered audit of one prominent claim --- the \emph{refinement
paradox}, in which more aggressive per-passage refinement reduces
answer faithfulness --- and use it to expose how
evaluation choices shape RAG conclusions.
We run \(>16{,}000\) pre-registered evaluations across SQuAD,
PubMedQA, HotpotQA, NaturalQuestions, TriviaQA, MS-MARCO, and QASPER
with five generators (Mistral-7B, Llama-3-8B, Qwen-2.5-7B,
Llama-3.3-70B, GPT-OSS-120B) and three faithfulness scorers
(DeBERTa-v3 NLI, roberta-large-mnli, and a Mistral-as-judge
RAGAS-style rubric).

Five findings emerge.
\textbf{(i)} The published \(n{=}30\) paradox magnitude collapses at
\(n{=}500{\times}5\) seeds: only one of five seeds reaches
\(p<0.05\) on the paired contrast and the pooled bootstrap 95\,\% CI
includes zero.
\textbf{(ii)} Faithfulness scoring is metric-fragile: under
DeBERTa-v3 the paradox magnitude is \(0.011\); under roberta-large-mnli
it is \(0.032\); under a RAGAS-style LLM-judge it is \(0.140\),
with pairwise Pearson \(r=0.18\)--\(0.67\).
\textbf{(iii)} A pre-registered matched-similarity intervention
(\(n{=}200\), \(\pm 0.02\) similarity tolerance) does not support
context coherence as a causal mediator at fixed mean retrieval
similarity (paired Wilcoxon \(p=0.628\), Cohen's \(d_z=-0.017\),
95\,\% bootstrap CI \([-0.022,+0.017]\)).
\textbf{(iv)} Retrieval threshold \(\tau\) does not transfer cleanly
across datasets: \(\tau\) tuned on SQuAD, PubMedQA, or NaturalQuestions
flags as leakage-prone (diagonal-vs-off-diagonal recovery gap \(>0.03\));
TriviaQA and HotpotQA do not.
\textbf{(v)} Head-to-head comparison of HCPC-v2 against CRAG, RAPTOR-2L,
and Self-RAG (\(n{=}1{,}600\) evaluations) shows no method dominates:
CRAG wins on HotpotQA short-answer multi-hop;
HCPC-v2 narrowly ties RAPTOR-2L on SQuAD;
Self-RAG underperforms in this matched harness
(\(0.57\) faithfulness, 39--43\,\% hallucination, p99 latency
\(45\)--\(47\)\,s vs.\ 3--5\,s for the others).

Despite these null and weakened claims, one structural finding
survives: at matched per-passage similarity, coherence-preserving
uninformative noise produces a smaller faithfulness drop than random
off-topic noise (slope \(-0.043\) vs.\ \(-0.069\)), suggesting
context coherence carries signal beyond similarity even when CCS
itself is not the causal mechanism.
We release the benchmark, scripts, pre-registrations, and per-query
CSVs so future RAG-faithfulness work can be audited the same way.
