\section{Released Benchmark}
\label{sec:benchmark}

The pre-registered audit in this paper is reproducible end-to-end
on free hardware. We release every artifact required to replicate
the claims and to extend the audit to new methods.

\paragraph{Released artifacts.}

\begin{itemize}
\itemsep -2pt
\item \textbf{Hugging Face Dataset.}
\href{https://huggingface.co/datasets/saketmgnt/context-coherence-bench}%
{\texttt{saketmgnt/context-coherence-bench}}.
Loadable via \verb|load_dataset("saketmgnt/context-coherence-bench")|.
Five configs:
\texttt{coherence\_paradox} (the SQuAD/Mistral $n{=}500{\times}5$ headline cell),
\texttt{adversarial\_drift}, \texttt{adversarial\_disjoint},
\texttt{adversarial\_contradict}, \texttt{adversarial\_control}.

\item \textbf{Permanent DOI.}
\href{https://doi.org/10.5281/zenodo.19757291}{10.5281/zenodo.19757291}
on Zenodo, with sha256 manifest, license, citation entry.

\item \textbf{Pip package.}
\verb|pip install context-coherence==0.2.0|. Pure-numpy core
implementation of CCS and the gating policy used by HCPC-v2.

\item \textbf{LangChain integration.}
\texttt{integrations/langchain/coherence\_gated\_retriever.py},
a drop-in \texttt{BaseRetriever} that exposes the CCS-gated
intervention. Status: integration is functional and tested; we
report negative results on its causal claim
(\S\ref{sec:causal_intervention}) but the diagnostic remains useful
for retrieval-time monitoring.

\item \textbf{Per-query CSVs.} Every per-query result table for every
fix in this paper is committed to the public repository at
\texttt{data/revision/fix\_NN/per\_query.csv}. Aggregations live at
\texttt{results/revision/fix\_NN/}. Pre-registrations are at
\texttt{experiments/fix\_NN\_log.md}.

\item \textbf{Kaggle T4\(\times\)2 reproduction notebooks.}
\texttt{notebooks/revision\_session1\_kaggle\_fresh.ipynb} (Fix 1 +
Fix 5 + Fix 11),
\texttt{notebooks/revision\_fix2\_kaggle\_t4x2\_fresh.ipynb} (Fix 2),
\texttt{notebooks/revision\_fix3\_4\_kaggle\_t4x2\_fresh.ipynb}
(Fix 3 + Fix 4),
\texttt{notebooks/revision\_fix5\_11\_kaggle\_t4x2\_fresh.ipynb}
(Fix 5 + Fix 11),
\texttt{notebooks/revision\_fix6\_kaggle\_t4x2\_fresh.ipynb} (Fix 6
including Self-RAG).
Total compute under 25 hours of free Kaggle T4\(\times\)2.

\item \textbf{Track-changes paper sources.} LaTeX sources for this
audit, with \texttt{\%~TRACKED-CHANGES:} markers indicating
exactly what changed from a prior published draft and why, are at
\texttt{ragpaper/sections/revision/}. The full revision book is at
\texttt{docs/revision/README.md}.
\end{itemize}

\paragraph{What's in the audit suite, in order.}

\begin{enumerate}
\itemsep -2pt
\item Construct matched-similarity HIGH/LOW CCS triples on a
target dataset (Fix 1 protocol). Pre-registered $\pm 0.02$ tolerance,
$n{=}200$ pairs.
\item Run the headline cell at $n{=}500$ across 5 seeds (Fix 2);
report per-seed paired Wilcoxon, pooled bootstrap CI, Wilson CI.
\item Re-score the headline cell with two additional faithfulness
scorers (Fix 3); report pairwise correlations.
\item Build the cross-dataset $\tau$-generalization matrix (Fix 4);
flag cells with diagonal-vs-off-diagonal recovery gap $> 0.03$.
\item Run coherence-preserving noise injection vs.\ random
noise injection (Fix 5); report regression slopes per condition.
\item Run head-to-head against published baselines on a matched
harness (Fix 6); report Pareto frontier.
\item For threshold-based methods, report cross-dataset numbers,
not in-distribution numbers.
\item Pre-register every $n$, every seed, every test
\emph{before} executing.
\end{enumerate}

We argue that any future RAG-faithfulness paper would benefit
from running this checklist before finalizing claims.
