\section{Limitations}
\label{sec:limitations}

\paragraph{Single-generator focus on the headline cell.}
The $n{=}500{\times}5$ scaling study (\S\ref{sec:scaling}) is on
SQuAD/Mistral-7B only. Multi-dataset numbers exist at $n{=}30$ and
$n{=}90$ for Mistral, Llama-3, and Qwen-2.5 from earlier rounds and
are released alongside this paper, but only the SQuAD/Mistral cell
has been re-run at proper power. A reviewer who wanted the same
audit on PubMedQA/Llama-3 (a more domain-mismatched setting) would
have to re-run that cell themselves. The released Kaggle notebooks
make this a 4--6 hour task on a free T4\(\times\)2.

\paragraph{Long-form QA out of scope.}
The audit is scoped to short-answer extractive QA. Long-form
generation evaluation (MS-MARCO, QASPER) is reported in earlier
work as not exhibiting the refinement paradox; our scaling study
does not extend to long-form. Long-form RAG faithfulness is an
open and harder evaluation problem
\citep{krishna2021hurdles}; we do not address it here.

\paragraph{Free 70B reproduction is budget-blocked.}
Independent reproduction of the frontier-scale (70B parameter)
faithfulness numbers via a non-Groq backend such as Together.ai
would require paid API access. Under the zero-dollar constraint
of this work that reproduction is blocked. The 70B/120B numbers
reported in earlier work via Groq's free tier (Llama-3.3-70B
SQuAD paradox $\approx 0.100$; GPT-OSS-120B SQuAD paradox
$\approx 0.030$ on $n{=}30$) stand as single-backend observations
and should be read with the understanding that we have not yet
reproduced them on a second 70B backend. This is disclosed, not
papered over.

\paragraph{RAGAS-judge self-bias risk.}
The RAGAS-style LLM-judge in our \S\ref{sec:multimetric}
multi-scorer audit uses a separate Mistral-7B instance with a
different prompt and a structured JSON-schema response. This
mitigates direct identity self-bias but not family-level bias: a
Mistral-7B judging a Mistral-7B answer may agree with the
generated answer's surface plausibility for reasons that have
nothing to do with faithfulness. A future version of this audit
using a frontier-scale third-party judge (Claude-3.5-Sonnet,
GPT-4o, Gemini-2-Pro) would resolve this. Doing so honestly costs
paid API tokens, which the zero-dollar constraint of this work
disallows.

\paragraph{Self-RAG harness fairness.}
\S\ref{sec:headtohead}'s Self-RAG numbers reflect Self-RAG run inside
a CRAG-style harness with a DeBERTa-v3 entailment scorer that does
not understand reflection tokens. The
``Self-RAG underperforms by $\sim 0.14$ faithfulness'' finding is
therefore harness-conditional. A more faithful Self-RAG comparison
would require modifying the entailment scorer to strip Self-RAG's
control tokens before scoring, or running Self-RAG on the
faithfulness benchmark its authors used. We report the
matched-harness number because it is the number a downstream
practitioner would observe; an aligned-harness number would
require additional engineering and is not in this paper's scope.

\paragraph{Causal intervention is single-domain.}
The Fix 1 matched-similarity intervention runs on SQuAD only,
$n{=}200$ matched pairs. The same intervention on PubMedQA or
HotpotQA is a natural extension that would either replicate or
contradict our null. We do not yet know which.

\paragraph{Single retrieval embedder for the scaling study.}
The scaling study (\S\ref{sec:scaling}) uses
\texttt{all-MiniLM-L6-v2} as the embedder. Earlier multi-retriever
work showed the paradox is alignment-mismatch-property:
the magnitude on SQuAD with stronger embedders (BGE-large,
GTE-large) is much smaller than with MiniLM. The collapse at scale
that we report is therefore consistent with what we would have
predicted from the multi-retriever results, but a full multi-
retriever multi-seed multi-metric study at $n{=}500$ remains future
work.
