\section{Methodology}
\label{sec:methodology}

\paragraph{Datasets.}
Seven question-answering corpora are used. SQuAD
\citep{rajpurkar2016squad}, PubMedQA \citep{jin2019pubmedqa},
HotpotQA distractor \citep{yang2018hotpotqa}, NaturalQuestions
\citep{kwiatkowski2019nq}, and TriviaQA \citep{joshi2017triviaqa}
form the short-answer extractive-QA core. MS-MARCO
\citep{nguyen2016msmarco} and QASPER \citep{dasigi2021qasper}
provide long-form retrieval evaluation contexts but are reported as
limitation-bounded scope (see \S\ref{sec:limitations}).

\paragraph{Generators.}
Five generators span the parameter range:
Mistral-7B-Instruct \citep{jiang2023mistral} and Llama-3-8B
\citep{meta2024llama3} via local Ollama, Qwen-2.5-7B-Instruct
\citep{qwen2024} also via Ollama, plus Llama-3.3-70B-Versatile
and GPT-OSS-120B via Groq's free API for frontier-scale checks.
Independent reproduction at 70B+ via Together.ai is
budget-blocked under the zero-dollar constraint of this work and
is disclosed as such (\S\ref{sec:limitations}).

\paragraph{Retrieval substrate.}
ChromaDB v0.4 \citep{chroma2023} with cosine similarity, indexed
over $1024$-token chunks (overlap $100$). The default embedder is
\texttt{sentence-transformers/all-MiniLM-L6-v2}
\citep{reimers2019sbert}. Multi-retriever robustness uses BGE-large,
E5-large, and GTE-large \citep{xiao2023bge,wang2022e5,li2023gte}.
Cross-encoder reranking (where required) uses
\texttt{ms-marco-MiniLM-L-6-v2}.

\paragraph{Faithfulness scorers.}
Three independent scorers are used in parallel:
\begin{itemize}
\itemsep -2pt
\item \(\phi_{\mathrm{DeBERTa}}\):
\texttt{cross-encoder/nli-deberta-v3-base} sentence-by-sentence
mean entailment \citep{he2021debertav3}.
\item \(\phi_{\mathrm{mnli}}\): \texttt{roberta-large-mnli} as a
zero-shot entailment proxy.
\item \(\phi_{\mathrm{ragas}}\): a Mistral-7B-as-judge instance
prompted with a RAGAS-style \citep{shahul2024ragas} faithfulness
rubric and a structured JSON-schema response, parsed defensively.
The judge is run as a \emph{separate} Mistral-7B process (different
context window, different instruction) from the system under test
to limit immediate self-bias; we discuss residual self-bias risk
in \S\ref{sec:multimetric}.
\end{itemize}
A faithfulness reading is reported as the per-query mean of one of
these scorers; hallucination rate is reported as the proportion of
queries for which the reading is below $0.5$, with Wilson score
95\,\% confidence intervals on that proportion.

\paragraph{Pre-registration log.}
Every experiment is pre-registered at
\texttt{experiments/fix\_NN\_log.md} with: hypothesis, sample size,
seed list, statistical test, effect-size threshold, and
honest-reporting commitment if the hypothesis is not supported. A
hypothesis is considered supported only if all three of the
following hold: (a) the pre-registered statistical test rejects at
$p<0.05$; (b) Cohen's $d_z>0.2$ for paired-difference experiments;
(c) the 10{,}000-resample bootstrap 95\,\% CI on the effect-size
estimate excludes zero. Every CSV reported in this paper is
committed at \texttt{data/revision/fix\_NN/per\_query.csv} and
every aggregated table at
\texttt{results/revision/fix\_NN/}. A reproducibility script
runs the full audit on a free Kaggle T4\(\times\)2 session in roughly
five hours; see \S\ref{sec:benchmark}.

\paragraph{Computational budget.}
All experiments run on free hardware: an Apple M4 Air for
Ollama-backed runs and a Kaggle T4\(\times\)2 session for the
8-bit Self-RAG comparison and any GPU-resident NLI scoring.
No paid APIs, no Groq quota burn beyond the free tier, no
Together.ai paid calls, and no GPT-4o-mini/Claude judge calls
are used. Total compute is well under 50 GPU-hours of free-tier
allocation across all experiments in this paper.
