\section{Related Work}
\label{sec:related}

\paragraph{Retrieval-augmented generation.}
Retrieval-augmented generation as a deployment pattern was
formalized by \citet{lewis2020rag} and rapidly extended through
fusion-in-decoder \citep{izacard2020fid}, end-to-end retrieval
training \citep{sachan2021endtoend}, and a growing collection of
post-retrieval refinement strategies. \citet{gao2023rag} provide a
recent survey. Our paper does not introduce a new RAG method; we
audit how existing methods are evaluated.

\paragraph{Retrieval-time interventions.}
The four published interventions we compare against
(\S\ref{sec:headtohead}) span the design space.
RECOMP \citep{xu2023recomp} compresses retrieved passages.
Self-RAG \citep{asai2024selfrag} delegates retrieve/skip decisions
to reflection tokens emitted by a fine-tuned generator.
CRAG \citep{yan2024crag} runs a per-passage relevance evaluator and
falls back to web search on low-confidence retrievals
(we substitute a cross-encoder for the unreleased T5 evaluator and
disable the web fallback for matched harness fairness).
RAPTOR \citep{sarthi2024raptor} builds a hierarchical summary tree
over the retrieval corpus offline.
Active-retrieval Forward-Looking Active REtrieval generation
\citep{jiang2023active} interleaves retrieval with generation.
None of these papers report cross-dataset $\tau$ generalization or
multi-scorer faithfulness; our audit is intended to make those
checks routine.

\paragraph{Faithfulness measurement.}
Sentence-level entailment via DeBERTa-v3 NLI
\citep{he2021debertav3} is the dominant choice in recent RAG-
faithfulness papers. Alternatives include FactScore
\citep{min2023factscore}, AlignScore \citep{zha2023alignscore},
and the RAGAS framework \citep{shahul2024ragas}, which uses an LLM
judge with an atomic-claim decomposition. We do not benchmark
faithfulness scorers in absolute terms; our \S\ref{sec:multimetric}
finding is that \emph{relative agreement} among popular scorers on
the same RAG outputs is much weaker than the field's reliance on a
single scorer would suggest.

\paragraph{Reproducibility and pre-registration in NLP.}
Pre-registration is uncommon in NLP relative to clinical or
psychological research, but a recent line of work argues for it
\citep{vandeven2020preregistration,ulmer2022preregister}.
\citet{rogers2023reviewing} document widespread variance in NLP
empirical claims under reproduction, particularly when small
samples and unreported tuning are involved. \citet{schaeffer2023emergent}
showed that an apparent emergent-abilities phenomenon dissolved under
proper-power evaluation, in a structurally similar episode to the
collapse we report in \S\ref{sec:scaling}.
Our audit operationalizes that recommendation for the specific
case of RAG faithfulness, with sample sizes, seeds, statistical
tests, and effect-size thresholds fixed in machine-checkable logs
before any generation runs.

\paragraph{Negative results in machine learning.}
Workshops and venues for negative results
\citep{negativeresults2024} have argued that disconfirming
empirical claims is a contribution. Our paper sits in that tradition
without being a pure negative-results paper: \S\ref{sec:coherence_noise}
preserves a structural positive finding even after the headline
claim collapses.
