\section{Scaling: The Paradox at Proper Sample Size}
\label{sec:scaling}

Published reports of the refinement paradox use $n{=}30$ queries
per condition. We re-run the SQuAD/Mistral-7B headline cell at
$n{=}500$ across five seeds (41--45), three conditions (baseline,
HCPC-v1, HCPC-v2), for $7{,}500$ paired evaluations.

\paragraph{Pooled headline.}
Pooling across all five seeds (\(n=2{,}500\) per condition) and
reporting bootstrap 95\,\% CIs for continuous metrics and Wilson
score 95\,\% CIs for the hallucination proportion:

\begin{table}[H]
\centering
\small
\caption{SQuAD/Mistral-7B at $n{=}500{\times}5$ seeds = $2{,}500$
queries per condition. Continuous metrics: 10{,}000-resample
bootstrap 95\,\% CI. Hallucination rate: Wilson score 95\,\% CI.}
\label{tab:scaling-pooled}
\begin{tabular}{lccccc}
\toprule
Condition & $n$ & Faith.\ ($\pm$ 95\,\% CI) &
Halluc (Wilson 95\,\% CI) & Mean retr.\ sim & Refine rate \\
\midrule
Baseline   & 2500 & $0.6609\;[0.6549, 0.6668]$ & $0.146\;[0.133, 0.160]$ & $0.5322$ & $0.000$ \\
HCPC-v1    & 2500 & $0.6503\;[0.6447, 0.6560]$ & $0.147\;[0.134, 0.162]$ & $0.5689$ & $0.000$ \\
HCPC-v2    & 2500 & $0.6612\;[0.6553, 0.6671]$ & $0.143\;[0.130, 0.157]$ & $0.5322$ & $0.705$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Per-seed paired contrasts.}
The paradox magnitude is the per-seed paired
$\Delta = \overline{\phi}_{\text{baseline}} - \overline{\phi}_{\text{HCPC-v1}}$.
Recovery is $\overline{\phi}_{\text{HCPC-v2}} - \overline{\phi}_{\text{HCPC-v1}}$.
Pre-registered analysis: paired Wilcoxon at $p<0.05$, alt=\textit{greater}.

\begin{table}[H]
\centering
\small
\caption{Per-seed paired Wilcoxon signed-rank tests at $n{=}500$.
Each row is one seed; significance is the unadjusted $p<0.05$
threshold. The pre-revision n=30 estimate of paradox magnitude was
$0.069 \pm 0.004$.}
\label{tab:scaling-perseed}
\begin{tabular}{cccccccc}
\toprule
Seed & $n$ & $\Delta$\textsubscript{baseline-v1} & $p$ & $d_z$
     & $\Delta$\textsubscript{v2-v1} & $p$ & $d_z$ \\
\midrule
41 & 500 & $0.0107$ & $0.083$ & $0.072$ & $0.0132$ & $\mathbf{0.025}$ & $0.088$ \\
42 & 500 & $0.0063$ & $0.234$ & $0.043$ & $0.0030$ & $0.504$ & $0.021$ \\
43 & 499 & $0.0078$ & $0.174$ & $0.055$ & $0.0059$ & $0.215$ & $0.042$ \\
44 & 500 & $0.0204$ & $\mathbf{0.008}$ & $0.134$ & $0.0191$ & $\mathbf{0.008}$ & $0.126$ \\
45 & 500 & $0.0073$ & $0.203$ & $0.048$ & $0.0125$ & $\mathbf{0.036}$ & $0.081$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Result.}
Per-seed paradox magnitudes are
$\{0.011, 0.006, 0.008, 0.020, 0.007\}$, with
mean $\bar{\Delta}=0.011$ and standard deviation $\sigma=0.006$.
Only seed 44 reaches $p<0.05$ on the paradox contrast. The pooled
bootstrap 95\,\% CI on the paradox contrast is
$[-0.005, +0.027]$, which contains zero.\footnote{Reported in
\texttt{results/revision/fix\_02/paired\_contrasts.csv}.}
This is a substantial downgrade from the published
$n{=}30$ estimate of $0.069$: the
$10\times$-larger sample yields a $\sim 6\times$-smaller magnitude
and a confidence interval consistent with no effect at all.

The recovery contrast (HCPC-v2 vs.\ HCPC-v1) is more robust:
3 of 5 seeds reach $p<0.05$, mean $\bar{\Delta}_{\mathrm{rec}} =
0.011$. This is not by itself a useful effect size given the
underlying paradox has collapsed --- if the baseline-vs-v1 contrast
is in the noise, the v2-vs-v1 ``recovery'' is recovering from the
same noise, not a real effect.

\paragraph{Interpretation.}
The published refinement-paradox magnitude on SQuAD/Mistral-7B is a
sample-size artifact at this scale. Combined with the causal
intervention null in \S\ref{sec:causal_intervention}, the empirical
case for an HCPC-v2-style intervention as a controlled
faithfulness probe is materially weaker than has been reported.
Section \S\ref{sec:multimetric} examines whether the choice of
faithfulness scorer changes this picture.
