Retrieval-augmented generation (RAG) rests on a comforting assumption:
retrieve better evidence and the model will answer more faithfully.
Across 6{,}050+ queries spanning SQuAD and PubMedQA, two generator
families (Mistral-7B and Llama-3-8B), and four retrieval configurations,
we show the assumption is not merely fragile — it can invert. Naive
post-retrieval refinement, the default in deployed rerank-and-compress
stacks, raises per-passage similarity while cutting SQuAD faithfulness
from 0.7995 to 0.6407 and introducing 10.0\% hallucination where there
was none. The culprit is not the retriever but a property of the
retrieved \emph{set}: when refinement splinters a broad passage into a
narrow high-similarity fragment, it strips the connective evidence the
generator was relying on, and the generator silently fills the gap with
parametric memory.

We name this failure mode the \textbf{refinement paradox} and argue that
faithfulness in RAG is governed by a set-level property we call
\textbf{context coherence}, not by per-passage relevance. Three
converging lines of evidence support the reframing. \emph{First}, a
selective refinement policy (HCPC-v2) that respects coherence recovers
SQuAD faithfulness to 0.7874 with 0\% hallucination while intervening on
only 16.7\% of queries, and is stable across a $7\times$ range of
refinement rates — restraint, not algorithmic sophistication, is the
active ingredient. \emph{Second}, in a domain-mismatched regime
(general-purpose encoder, biomedical corpus), RAG is strictly
\emph{worse} than zero-shot prompting: hallucination rises from 8.0\% to
20.0\% on Mistral and from 4.0\% to 10.0\% on Llama-3 — a prediction the
coherence account makes but the retrieval-quality account does not.
\emph{Third}, a variance decomposition shows chunk size — the parameter
that most directly shapes coherence — accounts for 90.1\% of
between-factor faithfulness variance on SQuAD ($F=19.72$, $p<0.001$,
Cohen's $d=1.41$); prompt engineering explains $<0.1\%$.

We distill the account into the Context Coherence Score (CCS), a
lightweight retrieval-time diagnostic computable from pairwise chunk
embedding similarities without running the generator. CCS cleanly
separates high-coherence retrieval sets (0\% hallucination) from
low-coherence sets that precede generation failures, giving practitioners
a deployment-time guard-rail that does not require answer-level
evaluation. Seven robustness checks confirm the result is not a single-run
artifact: the SQuAD paradox is $0.069 \pm 0.004$ across three random
seeds (signal/$\sigma = 17\times$); a one-way ANOVA over four prompt
templates yields $p = 0.83$ (no prompt main effect); HCPC
outperforms a hierarchical RAPTOR baseline on 2/3 datasets; and at
frontier scale Llama-$3.3$-$70$B exactly reproduces the SQuAD paradox
magnitude ($0.100$ vs $0.100$ at $7$B) while a $120$B GPT-OSS
generator still shows a positive $+0.030$ paradox — ruling out a
small-model artifact. Two
limitations qualify scope: the paradox does \emph{not} generalize to
long-form generation on QASPER or MS-MARCO (HCPC instead reduces
unsupported-claim rate by $67\%$ on the latter), and on SQuAD the
paradox magnitude is comparable to, not larger than, generic
retrieval-noise sensitivity. The broader message: RAG evaluation has
been optimizing the wrong quantity. We have been measuring how good
each passage looks on its own, when what the generator actually needs
is evidence that fits together.
