\section{Conclusion}
\label{sec:conclusion}

The central finding of this paper is uncomfortable, because it
contradicts a premise most RAG engineering quietly depends on: that
improving retrieval quality will improve answer faithfulness. Across
6{,}050+ queries spanning two datasets, two model families, and four
retrieval configurations, we show the premise is unreliable — and in
some regimes inverted. Naive post-retrieval refinement improves local
matching scores while cutting faithfulness by up to 15.9 percentage
points; domain-mismatched retrieval can make a RAG system \emph{less}
faithful than the same model with no retrieval at all.

Once we stop treating retrieval as a per-passage problem, the pattern
resolves. Faithfulness depends on whether the retrieved passages form
a mutually supportive evidence set, not merely on whether each passage
looks individually relevant. HCPC-v2 helps not because it is clever,
but because it is conservative: it intervenes only when two independent
signals both flag a fragmentation risk, protects the highest-ranked
passage, and restores contiguity whenever it does refine. CCS is the
same idea compressed into a number a practitioner can read off at
retrieval time — before any generation has happened, and before any
downstream harm.

Three findings have immediate practical value. First, chunk size is the
dominant design parameter for faithfulness among the factors tested
(90.1\% of between-factor variance on SQuAD), outweighing prompt
engineering by a factor of roughly $450\times$. Second, post-retrieval
refinement should be applied selectively based on context coherence,
not as a default post-processing step. Third, coherence should be
validated before deploying RAG in any specialized domain where the
encoder is not matched to the corpus.

\paragraph{What we did not settle, and where to push next.}
The paradox is well-characterized at the scale we ran, but load-bearing
questions remain open. Scaling is one: does a 70B+ generator, with more
parametric capacity to fall back on, amplify the failure mode or absorb
it? Encoder alignment is another: fitting a domain-matched encoder
(BioLinkBERT, PubMedBERT) to the PubMedQA corpus should test whether
the coherence account predicts the right repair. Multi-hop and
conversational settings are the cases where we expect coherence to
matter \emph{more}, not less, because the generator must stitch evidence
across reasoning steps; if the paradox does not appear there, the
account needs rethinking. Finally, CCS is used here as a diagnostic;
turning it into an active objective — a retriever that optimizes
set-level coherence rather than per-passage relevance — would move
coherence from something we measure to something we directly retrieve
for. We see that last direction as the most likely to produce a
qualitatively different class of RAG systems, and the most in need of
careful evaluation before anyone ships it.
