\section{Introduction}
\label{sec:intro}

A RAG pipeline looks, from the outside, like a safety feature. An LLM
without retrieval can hallucinate; an LLM with retrieval is anchored to
a document it can cite. That is the story the field tells about why we
build these systems, and the story behind most of the engineering
investment in them~\citep{lewis2020rag, gao2024retrievalaugmented}:
retrieve more relevant passages, and the generator will speak more
faithfully. The story underwrites deployments that matter — clinical
decision support, legal research, enterprise search — settings where a
confident but ungrounded answer can hurt someone.

This paper is a report on what we found when we stopped believing the
story.

\paragraph{Two quantities that have been quietly conflated.}
Retrieval quality is a property of \emph{individual passages}: each
chunk is scored against the query on its own. Faithfulness is a property
of the \emph{answer}, which emerges from a generator reasoning over the
concatenation of every passage the retriever handed it. Between these
two quantities lies a variable standard RAG evaluation does not look at
— the internal coherence of the retrieved context \emph{as a set}. When
passages are stitched into a single prompt, their mutual consistency is
what lets the generator treat them as one piece of evidence. When that
consistency breaks, the generator does not stop; it bridges the gap with
whatever its parameters remember. The citation looks fine. The answer
does not.

\paragraph{Why this matters in deployment.}
The gap between retrieval quality and faithfulness is not a laboratory
curiosity. In domain-specific deployments — biomedical QA, legal
research, financial analysis — a general-purpose encoder routinely
returns passages that are \emph{each} topical but \emph{together}
inconsistent. Under those conditions, our experiments show, RAG produces
more hallucination than the same model with no retrieval at all. The
safety feature becomes the failure mode.

\paragraph{Why existing methods do not catch it.}
The field has invested heavily in retrieval-side improvements:
cross-encoder re-ranking~\citep{nogueira2019passage}, context
compression~\citep{xu2023recomp}, self-reflective
retrieval~\citep{asai2024selfrag}. All three optimize passage-level
relevance or brevity. None measure — or preserve — the coherence of the
context the generator finally sees. Post-retrieval refinement, the
practice of splitting lower-scoring passages into higher-scoring
sub-chunks, is the clearest illustration: it makes every passage look
better while silently destroying the connective tissue between them.

\input{figures/overview}

\paragraph{What we claim, and what we do not.}
\emph{Faithfulness in RAG is governed by set-level coherence, not
passage-level relevance.} We name this mediating property \textbf{context
coherence}, and the central empirical consequence \textbf{the refinement
paradox}: aggressive post-retrieval refinement raises per-chunk
similarity while reducing answer faithfulness by up to 15.9 percentage
points and introducing hallucination where there was none. The
contribution of this paper is \emph{the failure mode and its mediating
variable}, not a new state-of-the-art system. HCPC-v2 and CCS
(introduced below) are instruments we use to test the coherence account;
if a better instrument replaces them tomorrow, the account stands or
falls independently. Figure~\ref{fig:overview} summarizes the study
design and the core finding.

\paragraph{Key findings at a glance.}
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \item Retrieval similarity $\neq$ faithfulness — the two can move in
        opposite directions.
  \item Context coherence is the missing variable that mediates the
        gap.
  \item Naive post-retrieval refinement causes hallucination by
        fragmenting otherwise coherent evidence.
  \item RAG with a domain-mismatched encoder can be strictly worse than
        zero-shot prompting.
  \item A cheap retrieval-time coherence check (CCS) recovers the
        expected behavior without retraining.
\end{itemize}

\paragraph{Contributions.}
We make four contributions, each of which is hard to explain within the
standard retrieval-quality framing of RAG and natural within the
coherence framing we propose.

\begin{enumerate}
  \item \textbf{Retrieval quality and answer faithfulness can move in
        opposite directions, by a wide margin.}
        On SQuAD, naive post-retrieval refinement improves per-chunk
        similarity while dropping faithfulness from 0.7995 to 0.6407 and
        introducing 10.0\% hallucination ($-$15.9~pp). Per-passage
        similarity and faithfulness are nearly \emph{decoupled} across
        the three refinement regimes we compare, falsifying the
        assumption that drives most RAG optimization.

  \item \textbf{RAG can be strictly worse than having no retrieval at
        all.} On PubMedQA, RAG with a general-domain encoder raises
        Mistral's hallucination rate from 8.0\% (zero-shot) to 20.0\%
        and Llama-3's from 4.0\% to 10.0\%. The coherence account
        predicts this inversion; standard retrieval framing does not.
        This is the paper's strongest empirical consequence: when
        retrieved evidence is topically adjacent but semantically
        inconsistent, parametric memory — at least internally
        consistent — is the safer input.

  \item \textbf{Selective, coherence-aware refinement eliminates the
        paradox with minimal intervention.}
        HCPC-v2 recovers SQuAD faithfulness to 0.7874 with 0\%
        hallucination by intervening on only 16.7\% of queries, and
        remains stable across a $7\times$ range of refinement rates
        (7\%--48\%). The design principle that does the work is
        restraint, not cleverness.

  \item \textbf{Chunk size dominates faithfulness variance by orders of
        magnitude over prompt strategy.}
        ANOVA marks chunk size as the only significant factor on SQuAD
        ($F = 19.72$, $p < 0.001$, $\eta^2 = 0.100$, Cohen's $d = 1.41$),
        accounting for 90.1\% of between-factor variance; prompt
        strategy explains $<0.1\%$. This is consistent with the
        coherence account — chunk size is the knob that most directly
        controls how much connective evidence survives retrieval — and
        it inverts the common engineering priority, which still has
        teams tuning prompts before fixing their chunking.
\end{enumerate}

We distill the reframing into \textbf{CCS}, a retrieval-time diagnostic
computed from pairwise chunk embedding similarities. CCS separates
coherent retrieval sets (0\% hallucination, CCS~$= 1.00$ on SQuAD) from
incoherent ones (20\% hallucination, CCS~$= 0.57$ on PubMedQA) without
running the generator — a computationally cheap guard-rail with a
clean scientific claim behind it: coherence, not relevance, is what
the generator is sensitive to.

\FloatBarrier
