\section{Experimental Design}
\label{sec:method}

This section describes the controlled design used to \emph{test} the
coherence hypothesis, not to propose a system. We treat retrieval
configuration as an independent variable whose effect on generator behavior
we wish to measure. Two instruments carry the causal argument: a selective
refinement probe (HCPC-v2) that perturbs context coherence in a targeted
way, and the Context Coherence Score (CCS, \S\ref{sec:ccs_def}) that
measures it without running the generator.

\subsection{A shared pipeline as experimental substrate}
\label{sec:shared}

\input{figures/pipeline}
All conditions share a single retrieval and generation substrate so that
any observed differences in faithfulness are attributable to the retrieval
regime under test; Figure~\ref{fig:pipeline} shows the common architecture.

\begin{itemize}
  \item \textbf{Embedding:}
        \texttt{all-MiniLM-L6-v2}~\citep{reimers2019sbert}, a 384-dimensional
        sentence encoder, with vectors stored in ChromaDB under cosine
        similarity.
  \item \textbf{Generation:} Mistral-7B~\citep{jiang2023mistral} and
        Llama-3-8B~\citep{llama3} served via Ollama at temperature~0.1.
        Using two generator families guards against model-specific
        explanations of the paradox.
  \item \textbf{Faithfulness scoring:}
        \texttt{cross-encoder/nli-deberta-v3-base}~\citep{he2021deberta}.
        Each answer sentence is classified against the concatenated
        retrieved context; the mean entailment probability is the
        faithfulness score in $[0,1]$. Responses below 0.50 are labeled
        hallucinated. The NLI scorer is held fixed across all conditions,
        so it functions as a constant measurement instrument rather than a
        tuned component.
\end{itemize}

The details above are \emph{experimental substrate}, not design claims.
None of our conclusions depend on the specific choice of encoder or
generator except where explicitly noted.

\subsection{Retrieval regimes as independent variables}

We compare four retrieval regimes that differ only in how the set of
passages reaching the generator is selected. They are designed to vary
context coherence while holding everything else constant.

\paragraph{Baseline (no refinement).}
Fixed-length chunks of 256, 512, or 1024 tokens with 10\% overlap; top-$k$
($k\in\{3,5\}$) by cosine similarity. This regime preserves whatever
coherence the document structure provides, since chunks are contiguous
spans.

\paragraph{Aggressive refinement (HCPC-v1).}
After top-$k$ retrieval, each chunk is rescored with a cross-encoder; if
\emph{either} cosine similarity $<0.45$ \emph{or} cross-encoder logit
$<-0.20$, the chunk is split into 256-token sub-chunks and the
highest-scoring fragment replaces the original. This OR-gated refinement is
the standard recipe in deployed rerank+compress RAG
stacks~\citep{xu2023recomp, nogueira2019passage}. We include it not as a
strawman but as a faithful instantiation of the prevailing
maximize-per-passage-relevance philosophy---it is precisely the regime the
refinement paradox indicts.

\paragraph{Selective refinement (HCPC-v2): a causal probe.}
HCPC-v2 is a controlled perturbation of HCPC-v1 designed to surgically
\emph{remove} its coherence-destroying failure modes while preserving its
ability to discard genuinely weak passages. If context coherence is the
causal mediator of faithfulness, this perturbation should recover
faithfulness even though retrieval similarity is unchanged. Concretely, it
modifies three behaviors of HCPC-v1:

\begin{enumerate}
  \item \textbf{AND-gated triggering} ($\tau_{\text{sim}}{=}0.45$,
        $\tau_{\text{ce}}{=}-0.20$): refinement fires only when both
        signals flag weakness, preventing over-triggering on adequate
        passages.
  \item \textbf{Rank protection}: the top-$N_{\text{prot}}{=}2$ ranked
        chunks are never refined, protecting the highest-evidence
        candidates from fragmentation.
  \item \textbf{Contiguity-preserving merge-back}: adjacent sub-chunks from
        the same parent are re-joined (up to $\sim$1024 tokens) when both
        pass the quality threshold, restoring connective evidence lost to
        splitting.
\end{enumerate}

\input{figures/architecture}
Figure~\ref{fig:architecture} illustrates the decision flow. We emphasize
that HCPC-v2 is not the paper's contribution. The choices of
$\tau_{\text{sim}}$, $\tau_{\text{ce}}$, and $N_{\text{prot}}$ are
engineering details that make the probe implementable; the scientific
question is whether \emph{any} coherence-preserving perturbation of the
refinement pipeline recovers faithfulness at fixed retrieval similarity.
Threshold sensitivity analysis (\S\ref{sec:results}) shows the recovery is
robust across a $7\times$ range of refinement rates---evidence that the
effect is driven by coherence, not threshold tuning.

\paragraph{Cross-encoder reranking (scoring-only control).}
A rerank-without-refine condition over-fetches $k_{\text{fetch}}{=}7$
candidates and returns the top $k{=}3$ by cross-encoder score, with no
splitting or merging. This isolates the effect of improved passage
\emph{selection} from the effect of passage \emph{fragmentation}, ruling
out alternative explanations of the paradox grounded in reranker behavior
alone.

\subsection{Why this design isolates context coherence}

The four regimes differ along exactly one dimension that matters for our
hypothesis: how aggressively the original contiguity of retrieved evidence
is broken. Baseline preserves it fully. Reranking preserves it while
reordering. HCPC-v1 destroys it indiscriminately. HCPC-v2 destroys it
selectively. If coherence is causally linked to faithfulness, faithfulness
should track this ordering---and we find that it does.

\input{figures/failure_case}
Figure~\ref{fig:failure_case} shows the mechanism concretely in a single
query: local sub-chunks score higher against the query while discarding the
very sentence the answer relied on.

\subsection{Measurement framework}

We report six quantities per condition:
\begin{enumerate}
  \item \textbf{NLI faithfulness} (mean entailment $\in[0,1]$).
  \item \textbf{Hallucination rate} (fraction $<0.50$).
  \item \textbf{Mean query--chunk similarity} (passage-level retrieval
        quality).
  \item \textbf{CCS} (Eq.~\ref{eq:ccs}; coherence of the retrieved set).
  \item \textbf{Refinement rate} (fraction of queries whose context was
        perturbed).
  \item \textbf{End-to-end latency} (for completeness, not a contribution).
\end{enumerate}

Statistical analysis uses one-way ANOVA for between-factor effects, with
effect sizes reported as $\eta^2$ and pairwise Cohen's $d$. Spearman's
$\rho$ is used for coherence--faithfulness correlations because retrieval
metrics are not normally distributed.

\FloatBarrier
