\section{Results}
\label{sec:results}

\subsection{Evidence for the refinement paradox}

Table~\ref{tab:hcpc_main} reports the central behavioral result. The
three regimes are matched on retrieval similarity to within 0.01, so the
faithfulness differences they produce must originate elsewhere. We argue
that the residual is context coherence.

\begin{table}[!htb]
\centering
\caption{HCPC head-to-head comparison (Mistral-7B, chunk = 1024, $k = 3$,
$n = 30$ per condition). The paradox is immediately visible: HCPC-v1
\emph{maintains or improves} retrieval similarity while \emph{degrading}
faithfulness and introducing hallucination.}
\label{tab:hcpc_main}
\small
\begin{tabular}{llccccc}
\toprule
Dataset & Condition & Faith. & Halluc. & Mean Sim & Refine\% & CCS \\
\midrule
SQuAD    & Baseline & 0.7995 & 0.0\%  & 0.634 & ---    & --- \\
SQuAD    & HCPC-v1  & 0.6407 & 10.0\% & 0.624 & 100.0\% & --- \\
SQuAD    & HCPC-v2  & 0.7874 & 0.0\%  & 0.634 & 16.7\% & 1.00 \\
\midrule
PubMedQA & Baseline & 0.5783 & 20.0\% & 0.653 & ---    & --- \\
PubMedQA & HCPC-v1  & 0.5808 & 26.7\% & 0.678 & 90.0\% & --- \\
PubMedQA & HCPC-v2  & 0.5893 & 20.0\% & 0.653 & 73.3\% & 0.57 \\
\bottomrule
\end{tabular}
\end{table}

The paradox manifests differently across datasets, revealing its
dependence on upstream retrieval quality. On SQuAD, where the
encoder--corpus alignment is strong, HCPC-v1 turns an already coherent
retrieval set into a worse answer context: faithfulness drops by 15.9
percentage points and hallucination rises from 0\% to 10\%. The
mechanism is over-intervention — the OR gate triggers on 100\% of
queries, replacing adequate passages with narrow fragments that break
the connective tissue between evidence segments. On PubMedQA, HCPC-v1
raises mean retrieval similarity from 0.653 to 0.678 — an improvement
under standard evaluation — yet hallucination rises from 20.0\% to
26.7\%. The co-occurrence confirms that higher per-chunk similarity can
coexist with lower answer quality when the intervention fragments
context coherence.

The following example from the SQuAD evaluation illustrates the mechanism
concretely.

\begin{quote}
\small
\textbf{Query:} ``What city did Super Bowl 50 take place in?''

\textbf{Baseline} (1024-token chunks, no refinement): The retrieved context
contains a coherent passage covering Super Bowl 50's venue, location, and
surrounding logistics. The generator answers: \emph{``Super Bowl 50 took
place in Santa Clara, California (specifically at Levi's Stadium in the San
Francisco Bay Area).''} Faithfulness: 0.9733.

\textbf{HCPC-v1} (naive refinement): All three chunks are flagged by the OR
gate and split into 256-token sub-chunks. The replacement fragments score
higher on query similarity ($\text{sim} = 0.589$) but the sub-chunk
containing the venue location is discarded in favor of one discussing the
game's branding as ``Super Bowl L.'' The generator responds: \emph{``I
cannot find this information in the provided context. The context only states
that Super Bowl 50 was a game to determine the champion\ldots''}
Faithfulness: 0.4730 (hallucinated).

\textbf{HCPC-v2} (selective refinement): The AND gate does not trigger (both
signals must agree), so the original coherent chunks are preserved.
The generator produces the same correct answer as baseline. Faithfulness:
0.9733.
\end{quote}

\noindent The example shows the paradox in microcosm: refinement raised
the per-chunk similarity score while removing the specific evidence the
generator needed. HCPC-v2 avoided the failure by not intervening at all.

A second example from PubMedQA illustrates the complementary failure
mode---coherence loss through terminology fragmentation rather than
evidence removal.

\begin{quote}
\small
\textbf{Query:} ``Does statin therapy reduce cardiovascular mortality in
diabetic patients?''

\textbf{Zero-shot} (no retrieval): The generator draws on parametric
knowledge and responds: \emph{``Yes, several large-scale trials including
the Heart Protection Study have shown that statin therapy significantly
reduces cardiovascular events in patients with diabetes.''} Faithfulness:
0.6340 (adequate; no hallucination).

\textbf{RAG baseline} (1024-token chunks, $k=3$): Three abstracts are
retrieved. Abstract~1 discusses ``HMG-CoA reductase inhibitors'' and
reports hazard ratios. Abstract~2 uses ``lipid-lowering agents'' and
reports odds ratios for ``major adverse cardiac events (MACE).''
Abstract~3 refers to ``atorvastatin'' specifically and reports
``all-cause mortality.'' Each abstract is individually relevant
($\text{sim} \geq 0.61$), but the three use non-overlapping terminology
for the same drug class, different outcome measures, and different
effect-size conventions. CCS~$= 0.54$. The generator produces:
\emph{``HMG-CoA reductase inhibitors reduce MACE, and atorvastatin
lowers all-cause mortality in diabetics, suggesting that lipid-lowering
agents are broadly protective.''} The answer stitches the three
terminologies into a composite claim that no single source supports;
the NLI scorer classifies it as unfaithful (0.4510) because the
synthesized assertion is not entailed by any retrieved passage.

\textbf{What coherence loss looks like here:} Unlike the Super Bowl case,
no evidence was \emph{removed}---all three abstracts reached the
generator intact. The failure is that the passages lack the
terminological bridges (e.g., ``statins, also known as HMG-CoA reductase
inhibitors'') that would signal to the generator that they describe the
same intervention. Absent these bridges, the generator treats the
passages as complementary evidence for a broader claim, producing a
hallucinated synthesis that each passage individually appears to support
but that the set collectively does not entail.
\end{quote}

\noindent This case demonstrates that coherence loss is not limited to the
fragment-replacement mechanism of aggressive refinement; it can originate
upstream, in encoder-corpus mismatch that assembles topically adjacent but
terminologically disjoint evidence. CCS flags the risk (0.54) before
generation; passage-level similarity does not.

\input{figures/paradox_plot}
Figure~\ref{fig:paradox} visualizes the disconnect between retrieval
similarity and faithfulness across all conditions.

\input{figures/qualitative_paradox}

\subsection{A targeted coherence-preserving perturbation restores faithfulness}

\input{figures/retrieval_and_tradeoff}
If coherence is the causal mediator, perturbations that preserve it should
restore faithfulness even when passage-level similarity is held fixed. The
HCPC-v2 probe tests exactly this prediction. It is not introduced as a
better algorithm; its role is to vary one quantity (coherence preservation)
while holding retrieval similarity constant.
Figure~\ref{fig:retrieval_comparison} compares faithfulness and hallucination
across all major retrieval conditions, and
Figure~\ref{fig:refinement_tradeoff} shows that the faithfulness recovery is
tightly coupled to the reduction in refinement rate.

On SQuAD, the selective probe recovers faithfulness to near-baseline
(0.7874 vs.\ 0.7995) and eliminates the hallucinations induced by
aggressive refinement, while intervening on only 16.7\% of queries (vs.\
100\% for HCPC-v1). On PubMedQA, the recovery is smaller (0.5783 to
0.5893): the residual failure cannot be repaired downstream because the
coherence loss originates upstream, in an encoder--corpus mismatch that
no refinement policy can correct. The asymmetry is itself informative
— it predicts, correctly, that the paradox is most severe in the regime
where passage-level similarity is healthy but evidence coherence is not.

\subsection{An upstream determinant of coherence dominates faithfulness variance}

Chunk size is not conventionally thought of as a coherence parameter, yet
it directly controls how much connective evidence a retrieved passage
contains. If coherence mediates faithfulness, chunk size should have an
outsized influence compared to parameters (like prompt wording) that do
not affect the evidence structure the generator reads. The variance
decomposition confirms this prediction. On SQuAD, one-way ANOVA yields $F = 19.72$
($p < 0.001$, $\eta^2 = 0.100$) for chunk size, compared to $F = 0.046$
($p = 0.830$, $\eta^2 < 0.001$) for prompt strategy. Among the three
configuration factors tested (chunk size, top-$k$, and prompt strategy),
chunk size accounts for 90.1\% of the between-factor variance
(Appendix~\ref{app:variance_model}). The 256-to-1024 token effect
corresponds to Cohen's $d = 1.41$---a large effect by conventional
standards---reflecting that larger chunks preserve complete sentences,
logical transitions, and redundant evidence that collectively reduce
hallucination risk.

This result has an immediate practical implication: it inverts the
common engineering priority of optimizing prompt templates before
chunking strategy. Among the parameters we tested, the faithfulness
return from chunk-size selection is roughly 450 times larger than from
prompt style
($\eta^2_{\text{chunk}} / \eta^2_{\text{prompt}} = 0.100 / 0.0002$).

\subsection{Robustness: the effect is not a threshold artifact}

A natural alternative explanation for the selective-probe result is that
we tuned thresholds until the paradox vanished. The Phase~5 sweep rules
this out. Table~\ref{tab:threshold} reports representative SQuAD
operating points across the sweep.

\begin{table}[!htb]
\centering
\caption{Representative threshold configurations on SQuAD. Faithfulness
remains stable across refinement rates from 7\% to 48\%, degrading only
when refinement becomes aggressive enough to erode context coherence. This
broad plateau means HCPC-v2 does not require precise tuning for deployment.}
\label{tab:threshold}
\small
\begin{tabular}{ccccc}
\toprule
$\tau_{\text{sim}}$ & $\tau_{\text{ce}}$ & Faith. & Halluc. & Refine\% \\
\midrule
0.60 & $-0.05$ & 0.7810 & 3.3\% & 7\%  \\
0.55 & $-0.10$ & 0.7815 & 3.3\% & 17\% \\
0.45 & $-0.20$ & 0.7823 & 3.3\% & 33\% \\
0.40 & $-0.30$ & 0.7809 & 3.3\% & 48\% \\
0.35 & $-0.40$ & 0.7506 & 3.3\% & 68\% \\
\bottomrule
\end{tabular}
\end{table}

\input{figures/threshold_plateau}
Figure~\ref{fig:threshold_plateau} shows a broad performance plateau
followed by a sharp decline once refinement exceeds approximately 50\% of
queries. The plateau behavior has a coherence-based explanation. Below 50\%
refinement, the method is modifying passages that are genuinely weak, and
the merge-back step restores contiguity in most cases. Above 50\%, the
method begins refining passages that were adequate, and the cumulative
fragmentation degrades the context faster than merge-back can repair it.

\subsection{A stronger test: when retrieval itself is harmful}

The contrast between SQuAD and PubMedQA is summarized in
Figure~\ref{fig:dataset_comparison}.

PubMedQA is not simply harder in the conventional sense. Its mean
retrieval similarity is comparable to SQuAD's, yet its faithfulness is
consistently lower and its hallucination rate higher. The difference
lies in context coherence: the retrieved biomedical abstracts are
topically adjacent but not mutually supportive — they use different
terminology for related concepts, present contradictory evidence from
different studies, and lack the narrative continuity of Wikipedia
passages. This is exactly the setting where CCS captures what raw
similarity scores miss.

The zero-shot comparison makes the deployment consequence stark.
Table~\ref{tab:zeroshot} shows that RAG helps on SQuAD but hurts on
PubMedQA, and Figure~\ref{fig:zeroshot} visualizes the divergence.

\begin{table}[!htb]
\centering
\caption{Zero-shot versus best RAG configuration. On SQuAD, retrieval
reduces hallucination as expected. On PubMedQA, retrieval with a
general-domain encoder \emph{increases} hallucination relative to answering
without any retrieved context---a warning for domain-specific RAG
deployment where the encoder is not matched to the corpus.}
\label{tab:zeroshot}
\small
\begin{tabular}{llcccc}
\toprule
Model & Dataset & ZS Faith. & ZS Halluc. & RAG Faith. & RAG Halluc. \\
\midrule
Mistral & SQuAD    & 0.6895 & 5.0\%  & 0.7775 & 3.0\% \\
Mistral & PubMedQA & 0.6118 & 8.0\%  & 0.6004 & 20.0\% \\
Llama-3 & SQuAD    & 0.6967 & 13.0\% & 0.7752 & 0.0\% \\
Llama-3 & PubMedQA & 0.6110 & 4.0\%  & 0.6233 & 10.0\% \\
\bottomrule
\end{tabular}
\end{table}

\input{figures/dataset_and_zeroshot}

On PubMedQA, Mistral's hallucination rate rises from 8.0\% (zero-shot) to
20.0\% (best RAG)---a $2.5\times$ increase---and Llama-3's from 4.0\% to
10.0\%. This is the strongest prediction of the coherence framework: when
the retrieved set is topically adjacent but not mutually supportive, the
generator is worse off than with no retrieval at all. Parametric memory,
while less specific, is at least internally consistent; a low-coherence
retrieved context is not. The effect is a direct behavioral signature of
coherence-driven failure, not an artifact of any particular pipeline
choice, and it is predicted \emph{a priori} by the framing we propose.

\FloatBarrier
