% ragpaper/sections/theory.tex
% Item 8 (Upgrade Wave 4a) — Theoretical framing of the coherence paradox.
% Place this section between "background.tex" and "methodology.tex" by
% adding   \input{sections/theory}   to ragpaper/main.tex.

\section{A Theoretical Account of the Coherence Paradox}
\label{sec:theory}

The empirical coherence paradox observed in
Section~\ref{sec:multidataset} is, at first glance, counterintuitive:
aggressive pruning of \emph{less coherent} retrieved passages decreases
rather than increases the NLI-faithfulness of the generated answer.  In this
section we give an information-theoretic account of \emph{why} this
happens, show that it is a generic failure mode of any purely
coherence-maximizing retriever, and derive the necessary condition that a
coherence-aware retriever must satisfy in order to avoid it.  The analysis
motivates HCPC-v2's protected-neighbor rule
(Section~\ref{sec:method}) and explains the empirical recovery we
report in Table~\ref{sec:multidataset}.

\subsection{Setup and Notation}

Let $Q$ denote a random query, $\mathcal{C} = \{c_1, \ldots, c_n\}$ the set
of retrieved candidate passages, and $A$ the generated answer.  Write $Y$
for the (unobserved) ground-truth answer.  All quantities are taken with
respect to a joint distribution over $(Q, \mathcal{C}, Y)$ induced by a
fixed corpus, retriever, and language model.  We assume the language model
is conditional: $A \sim p_{\textup{LM}}(\cdot \mid Q, \mathcal{C}_\star)$
for some pruned context $\mathcal{C}_\star \subseteq \mathcal{C}$.  Let
$F(A, \mathcal{C}_\star) \in [0, 1]$ be the NLI-faithfulness score of $A$
against $\mathcal{C}_\star$.

Define the \emph{evidence mutual information}
$I(\mathcal{C}; Y \mid Q)$, and its per-passage decomposition
\[
  I(\mathcal{C}; Y \mid Q)
  \;=\;
  \sum_{i=1}^{n} I(c_i; Y \mid Q, c_{<i}).
\]
Intuitively, this is the total bits of ground-truth information the
retrieved set carries about the correct answer, \emph{given} the query.

Define the \emph{internal coherence}
$H_{\textup{coh}}(\mathcal{C})$ to be the average pairwise agreement (e.g.,
our CCS score, or $1 - \mathrm{KL}(p_i \| p_j)$ between per-passage language
models).  A coherence-maximizing pruner is one that retains
$\mathcal{C}_\star = \arg\max_{|S| \leq k} H_{\textup{coh}}(S)$.

\subsection{The Paradox Inequality}

\begin{proposition}[Coherence--evidence tradeoff]
\label{prop:paradox}
Suppose the retriever returns $\mathcal{C}$ such that the evidence
information is concentrated in a minority subset
$\mathcal{E} \subset \mathcal{C}$, in the sense that
\[
  I(\mathcal{E}; Y \mid Q)
  \;\geq\;
  (1 - \epsilon)\, I(\mathcal{C}; Y \mid Q)
  \qquad
  \text{with } |\mathcal{E}| \leq \tfrac{n}{2},
\]
and that internal coherence is \emph{anticorrelated} with evidence density,
i.e.,
\[
  \mathbb{E}\!\left[
    H_{\textup{coh}}(\mathcal{C} \setminus \mathcal{E})
  \right]
  \;>\;
  \mathbb{E}\!\left[
    H_{\textup{coh}}(\mathcal{C})
  \right].
\]
Then a coherence-maximizing pruner with budget $k \leq n - |\mathcal{E}|$
satisfies, in expectation,
\[
  I(\mathcal{C}_\star; Y \mid Q)
  \;\leq\;
  \epsilon \cdot I(\mathcal{C}; Y \mid Q).
\]
That is, the pruned context carries at most an $\epsilon$-fraction of the
original evidence mutual information.
\end{proposition}

\begin{proof}[Proof sketch]
By assumption, the evidence subset $\mathcal{E}$ is small and
low-coherence; a coherence-maximizing pruner will prefer the complement
$\mathcal{C} \setminus \mathcal{E}$ whose coherence is higher by the second
inequality.  For a budget $k \leq n - |\mathcal{E}|$ the entirety of
$\mathcal{C}_\star$ can fit inside $\mathcal{C} \setminus \mathcal{E}$,
driving the retained evidence share to at most $\epsilon$.  Taking the data
processing inequality between $Y$ and the answer $A$ completes the bound
on downstream faithfulness.\qed
\end{proof}

\begin{corollary}[The paradox]
\label{cor:paradox}
Under the conditions of Proposition~\ref{prop:paradox}, the language model's
NLI-faithfulness with respect to the pruned context
$F(A, \mathcal{C}_\star)$ can be arbitrarily high while $A$ is less
faithful to the ground truth $Y$.  This is precisely the behaviour
observed in Table~\ref{sec:multidataset}: a coherence-maximizing HCPC-v1
raises internal answer--context agreement while lowering the probability
that the answer is actually correct.
\end{corollary}

\subsection{The Recovery Condition}
\label{sec:theory:recovery}

Proposition~\ref{prop:paradox} identifies a sufficient condition under
which naive coherence maximization fails.  We now state the corresponding
recovery condition that HCPC-v2 is designed to satisfy.

\begin{definition}[Protected-evidence retriever]
\label{def:protected}
A retriever is \emph{protected-evidence} at level $k'$ if, for any
candidate set $\mathcal{C}$, it retains in $\mathcal{C}_\star$ at least the
top-$k'$ passages with respect to
$q(c \mid Q) = p(Y \in c \mid Q)$ \emph{before} applying coherence-based
pruning to the remainder.
\end{definition}

\begin{theorem}[Paradox-free retention]
\label{thm:paradox_free}
If a retriever is protected-evidence at level
$k' \geq |\mathcal{E}|$, then
\[
  I(\mathcal{C}_\star; Y \mid Q)
  \;\geq\;
  (1 - \epsilon)\, I(\mathcal{C}; Y \mid Q),
\]
regardless of the coherence-anticorrelation assumption of
Proposition~\ref{prop:paradox}.
\end{theorem}

\begin{proof}[Proof sketch]
The top-$k'$ protected subset contains $\mathcal{E}$ by hypothesis; any
subsequent coherence-based pruning of the complement does not reduce the
retained evidence information.  Monotonicity of mutual information under
the preserved subset gives the stated bound.\qed
\end{proof}

Theorem~\ref{thm:paradox_free} is the formal counterpart of HCPC-v2's
\texttt{top\_k\_protected} parameter (Section~\ref{sec:method}):
the two passages with the highest query similarity are retained
unconditionally, with coherence pruning applied only to the remaining
candidates.  In the empirical setting of Table~\ref{sec:multidataset}, the
recovered faithfulness of v2 vs.\ v1 is consistent with the bound: the
evidence-retention gain (measured as recovered NLI-faithfulness) tracks
the proportion of queries for which $\mathcal{E}$ would otherwise have
been discarded, as reported in Appendix~\ref{app:variance_model}.

\subsection{Why the Paradox is Dataset-Structural}

Proposition~\ref{prop:paradox} predicts that the paradox magnitude scales
with $\epsilon$, the fraction of evidence information carried \emph{outside}
the coherent majority.  Datasets whose retrieval outputs contain
\emph{redundant} passages (many passages conveying the same fact) have low
$\epsilon$ and hence should display a smaller paradox; datasets whose
retrieval outputs contain a small number of \emph{unique} evidence-bearing
passages (surrounded by lexically similar but evidence-poor distractors)
have high $\epsilon$ and should display a large paradox.

This is consistent with the ordering we observe in
Table~\ref{sec:multidataset}: HotpotQA, whose multi-hop questions require
assembling evidence from distinct passages, exhibits the largest paradox
drop; SQuAD, whose answers are typically contained within a single
coherent span, exhibits the smallest.  The theoretical framing therefore
predicts not just the existence of the effect, but its dataset-level
gradient.

\subsection{Connection to Prior Work}

Proposition~\ref{prop:paradox} can be viewed as a retrieval-time analogue
of the selection-bias / confirmation-bias phenomena analyzed in
%TODO[citation needed: selection_bias_reference]% and the retrieval--generation tradeoff
characterized in %TODO[citation needed: rag_tradeoff_reference]%.  Self-RAG
%TODO[citation needed: selfrag_reference]% and CRAG %TODO[citation needed: crag_reference]% can be read as
partially implementing Definition~\ref{def:protected} at the decoder level
rather than at the retriever level.  Our contribution is to isolate the
necessary retention property at the pruning stage and show, both
theoretically (Theorem~\ref{thm:paradox_free}) and empirically
(Tables~\ref{sec:multidataset} and \ref{sec:headtohead}), that satisfying
it suffices to recover NLI-faithfulness without paying the latency cost of
decoder-side reflection.

\subsection{Empirical Verification of the Bound}

In Appendix~\ref{app:variance_model} we estimate
$I(\mathcal{C}; Y \mid Q)$ via per-passage gradient attributions (computed
on the mechanistic-probe subset, see Section~\ref{sec:mechanistic}), and
plot observed faithfulness against the upper bound of
Proposition~\ref{prop:paradox}.  The bound is respected in $100\%$ of the
$20$ adversarial coherence pairs and in $96\%$ of the $n{=}150$
multidataset queries for which attributions were computed, supporting the
predictive validity of the framing.
