\section{Artifact Manifest}
\label{app:manifest}

% POLISH FIX 4: Added a harness-mismatched Self-RAG appendix table so the main
% comparison stays like-for-like while preserving artifact traceability.
The double-blind supplement includes repository-local source tables and
result files. Public links are anonymized for review; code and data will be
released upon acceptance.

\begin{table}[H]
\centering
\small
\caption{Main evidence sources used by the paper. Paths are repository-local
inside the anonymized artifact.}
\label{tab:source_manifest}
\begin{tabular}{ll}
\toprule
Claim area & Source files \\
\midrule
Matched CCS test & \verb|results/revision/fix_01/paired_wilcoxon.csv| \\
Scaled pilot cell & \verb|results/revision/fix_02/headline_table.csv| \\
Multi-metric scoring & \verb|results/revision/fix_03/table1_multimetric.csv| \\
Human calibration & \verb|data/revision/fix_03/human_eval_*.csv| \\
Threshold transfer & \verb|results/revision/fix_04/tau_transfer_matrix.csv| \\
Noise separation & \verb|results/revision/fix_05/slope_response.csv| \\
Full Self-RAG audit & \verb|data/revision/fix_06/per_query_full_selfrag.csv| \\
Baseline summary & \verb|results/revision/fix_06/h2h_summary_full_selfrag.csv| \\
RAPTOR cost table & \verb|results/revision/fix_11/raptor_full_table.csv| \\
\bottomrule
\end{tabular}
\end{table}

\section{Human-Evaluation Details}
\label{app:human_eval}

The human calibration uses 99 examples sampled from the fixed-generation
multi-metric cell. Two raters label each answer as \verb|supported|,
\verb|partially_supported|, or \verb|unsupported| using only the provided
retrieved passages. Disagreements are resolved in an adjudicated CSV.

\begin{table}[H]
\centering
\small
\caption{Human calibration label distribution after adjudication.}
\label{tab:human_label_distribution}
\begin{tabular}{lrr}
\toprule
Label & Count & Proportion \\
\midrule
Supported & 75 & 0.758 \\
Partially supported & 20 & 0.202 \\
Unsupported & 4 & 0.040 \\
\bottomrule
\end{tabular}
\end{table}

Raw agreement is 0.919 and Cohen's \(\kappa=0.774\). The correlation table in
the main paper is computed from adjudicated ordinal labels
(\verb|supported| \(=1\), \verb|partially_supported| \(=0.5\),
\verb|unsupported| \(=0\)). Kendall \(\tau_b\) values are 0.114 for DeBERTa,
0.311 for the second NLI scorer, and 0.421 for the RAGAS-style judge.

\paragraph{Annotation instructions.}
Raters judge answer support by the retrieved passages only. Outside knowledge
is disallowed. If the context supports the main answer but the answer adds an
unsupported qualifier or side detail, raters use \verb|partially_supported|.
If the answer says the information is unavailable, raters mark it supported
only when the provided context really omits the answer.

\section{Self-RAG Artifact Traceability}
\label{app:selfrag}

% POLISH FIX 4: Added a dedicated harness-mismatch appendix table so Self-RAG
% remains traceable without visually dominating the main matched-harness audit.
The full Self-RAG package has been imported into repository-local files, so the
paper does not depend on a path outside the anonymized artifact. The verified
import contains the expected 1,600 per-query rows and 8 summary rows.

\begin{table}[H]
\centering
\tiny
\setlength{\tabcolsep}{2pt}
\caption{Full Self-RAG artifact checksums.}
\label{tab:selfrag_checksums}
\begin{tabular}{p{0.43\linewidth}rp{0.42\linewidth}}
\toprule
File & Rows & SHA-256 \\
\midrule
\texttt{data/revision/fix\_06/per\_query\_full\_selfrag.csv} & 1600 &
\texttt{aaf00af2f186a3d90c619c3885df2627e237580c54ac5837be31443f1360a5af} \\
\texttt{results/revision/fix\_06/h2h\_summary\_full\_selfrag.csv} & 8 &
\texttt{5d325428d24c7d62b4fa22e30108edec1f5842d862536fc75b5f56192dc82055} \\
\texttt{results/revision/fix\_06/summary\_full\_selfrag.md} & -- &
\texttt{cba4cadcd97b2705c6aaef7893816e7d2bdc1ce8f2ee67d92dce1fe537ba1574} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Harness-Mismatched Baseline: Self-RAG}
\label{app:selfrag_harness}

Self-RAG is included for artifact traceability, but it is not a like-for-like
matched-harness baseline. The run uses the released fine-tuned Llama-2-7B
Self-RAG checkpoint with 8-bit loading, while CRAG, HCPC-v2, and RAPTOR-2L use
the Mistral-backed short-answer harness. Self-RAG is also trained for
reflective decoding, whereas the entailment-style scorer used in the audit
does not separately reward reflection tokens or task-specific citation
behavior. These numbers should therefore be read as harness-conditional, not
as a global comparison against Self-RAG.

\begin{table}[H]
\centering
\small
\caption{Harness-mismatched Self-RAG baseline. Latency is milliseconds.}
\label{tab:selfrag_harness}
\begin{tabular}{lrrrr}
\toprule
Dataset & Faith & Halluc. & Mean lat. & p99 lat. \\
\midrule
SQuAD & 0.574 & 0.390 & 7542 & 44800 \\
HotpotQA & 0.574 & 0.430 & 6158 & 46965 \\
\bottomrule
\end{tabular}
\end{table}

\section{Reproducibility Commands}
\label{app:commands}

All revision runs are designed for free local or Kaggle compute. The exact
runbook is included in \verb|docs/revision/runbook.md|. Representative
commands are:

\begin{verbatim}
python3 experiments/fix_01_causal_matched_pairs.py
python3 experiments/fix_02_scaled_headline_n500.py
python3 experiments/fix_03_multimetric_faithfulness.py
python3 experiments/analyze_fix03_human_eval.py
python3 experiments/fix_04_tau_generalization.py
python3 experiments/fix_05_coherence_preserving_noise.py
python3 experiments/fix_06_baseline_h2h_pareto.py
python3 experiments/fix_11_raptor_full_table.py
\end{verbatim}

\section{Anonymous Release Placeholders}
\label{app:release}

During double-blind review, public author-identifying links are replaced with
placeholders such as [anonymized repository], [anonymized dataset], and
[anonymized DOI]. Post-review sources can restore the public repository,
dataset, package, and DOI links after review.
