\section{Introduction}
\label{sec:introduction}

% POLISH FIX 1: Moved the fixed-generation multi-scorer result to the opening
% frame so the paper leads with its strongest positive audit finding.
RAG evaluation has a familiar shortcut: a retrieval score improves, one
automatic faithfulness metric moves in the desired direction, and the paper
infers that answer support has improved. Our audit finds that the scorer can
be the fragile link. On the same 7,500 generated answers, the same intervention
has effect sizes 0.011, 0.032, and 0.140 under DeBERTa, a second NLI scorer,
and a local RAGAS-style judge. In a 99-example two-rater calibration
(\(\kappa=0.774\)), DeBERTa has weak rank alignment with adjudicated support
labels, while the second NLI scorer and RAGAS-style judge are only moderately
aligned; their confidence intervals overlap, so the ordering is suggestive
rather than definitive. The durable finding is not that one scorer is best,
but that RAG faithfulness claims are not identified by a single observer.

This paper audits that gap. A generator receives a context set, not a list of
independent passage scores, and a faithfulness conclusion depends on the
observer used to score support. We use the refinement paradox as a stress
test: the pilot finding that more aggressive per-passage refinement can raise
retrieval similarity while lowering measured answer faithfulness. The goal is
not to promote that phenomenon as large or stable. It is to ask a sharper
methodological question:

\begin{quote}
\emph{Which RAG faithfulness claims remain identifiable once local retrieval
similarity, context-set structure, evaluator choice, human calibration,
threshold transfer, and cost are separated?}
\end{quote}

The answer is sobering. The strongest intuitive story -- that context
coherence explains faithfulness failures -- fails under a direct control. We
construct HIGH- and LOW-CCS context sets matched on mean query-passage
similarity; the coherence manipulation is large, but faithfulness does not
improve. This failure is the point. It shows how a plausible retrieval-time
quantity can look meaningful in aggregate tables and then lose its apparent
effect once the confounder is isolated.

The wider audit reaches the same conclusion from other angles. The pilot
DeBERTa effect becomes small and seed-sensitive at \(n=500\times5\). The same
generations imply different intervention magnitudes under different scorers.
A small two-rater human calibration clarifies that the scorer disagreement is
not cosmetic: no automatic scorer has strong human alignment in this sample,
although the RAGAS-style judge has the highest rank correlation. Thresholds
tuned on one dataset transfer unevenly. Baseline comparisons change when p99
latency and indexing cost are included.

% POLISH FIX 6: Promoted the minimum reporting standard from the discussion
% into the introduction so reviewers see the reusable recommendation early.
\paragraph{Recommended minimum reporting standard for RAG faithfulness claims.}
\begin{enumerate}[leftmargin=1.8em,itemsep=1pt,topsep=2pt]
  \item Paired query-level contrasts with bootstrap intervals.
  \item Local query-passage similarity and at least one set-level context statistic.
  \item Multiple faithfulness scorers.
  \item A small human calibration whenever automated metrics disagree.
  \item Threshold transfer across datasets.
  \item Hallucination rates with Wilson intervals.
  \item Mean latency, p99 latency, and offline indexing cost.
\end{enumerate}

\paragraph{Contributions.}
\begin{itemize}[leftmargin=1.6em,itemsep=2pt,topsep=2pt]
  \item We show that faithfulness scoring is metric-fragile on fixed
  generations, and calibrate the disagreement with a 99-example two-rater
  human audit.
  \item We provide a pre-registered scale audit of a pilot refinement-paradox
  cell, replacing a small-sample headline with \(n=500\times5\) seed evidence.
  \item We falsify the strong CCS interpretation with a matched-similarity
  intervention, while preserving a narrower diagnostic role via a
  coherent-versus-random noise test.
  \item We audit threshold transfer and cost-aware baselines, and document the
  Self-RAG artifact separately as a harness-mismatched appendix baseline.
  \item We package the checks as \method{}, a reusable protocol for auditing
  RAG faithfulness claims. Code and data are anonymized in the supplementary
  material and will be released upon acceptance.
\end{itemize}

The paper's thesis is deliberately conservative: RAG faithfulness claims are
under-identified unless retrieval similarity, context-set structure, evaluator
choice, human calibration, threshold transfer, and cost are reported separately.
