\section{Limitations and Ethics}
\label{sec:limitations_ethics}

% POLISH FIX 4/P2-B: Reframed Self-RAG as harness-mismatched and clarified
% that earlier large-model observations are not central revision evidence.
\paragraph{Short-answer scope.}
The main evidence concerns short-answer extractive QA. It should not be used
to claim the same behavior in long-form synthesis, agentic workflows, or
domain-specific retrieval without additional controls.

\paragraph{Human calibration and judge bias.}
The human evaluation has 99 examples, two raters, and adjudicated labels: it
calibrates scorer disagreement in the audited cell, but does not definitively
validate any faithfulness metric. The RAGAS-style judge is local and
zero-dollar, but may share family-level biases with the generator; the stronger
conclusion is metric fragility, not that the judge is universally correct.

\paragraph{CCS and threshold scope.}
The matched-similarity test is a single-domain causal test on SQuAD/Mistral.
It rejects the strong CCS interpretation in that setting; it does not exhaust
all possible context-structure hypotheses. Threshold transfer is also uneven,
so thresholded policies should report off-diagonal behavior.

\paragraph{Self-RAG harness fairness.}
Self-RAG is a harness-mismatched appendix baseline because it uses a different
backbone, quantization path, and reflective decoding objective than the
Mistral-backed methods in the main table. Its measured behavior should not be
read as a global statement about Self-RAG.

\paragraph{Large-model reproduction.}
The pilot's 70B and 120B observations are not re-verified under the zero-dollar
revision constraint and therefore are not treated as central evidence; they
are not contradicted by the audit.

\paragraph{Ethics.}
The audit is intended to reduce overclaims about faithful RAG systems. The
main risk is misuse: a single metric, a single threshold, or one favorable
baseline row could be cherry-picked to market a system as grounded. We
therefore emphasize per-query logs, multiple metrics, human calibration,
negative results, and cost-aware reporting.
