\section{Pre-Registered Audit Protocol: \method{}}
\label{sec:protocol}

The experiments are a controlled falsification suite. Each test isolates a
claim that can look convincing in a conventional aggregate RAG table, then
asks whether the claim survives once the relevant confounder is separated.

\paragraph{Pre-registration and statistics.}
Sample sizes, seeds, primary statistics, and decision rules were frozen before
execution. Paired continuous contrasts use Wilcoxon signed-rank tests and
Cohen's \(d_z\). Means use 10,000-resample bootstrap confidence intervals.
Hallucination rates use Wilson intervals. Thresholds are frozen except in the
threshold-transfer experiment, where tuning and evaluation datasets are
explicitly separated. No paid API calls are used in the revision results.

\paragraph{Datasets, generators, and scorers.}
The core revision evidence uses short-answer extractive QA, primarily SQuAD,
PubMedQA, HotpotQA, NaturalQuestions, and TriviaQA
\citep{rajpurkar2016squad,jin2019pubmedqa,yang2018hotpotqa,kwiatkowski2019nq,joshi2017triviaqa}.
The main scaled cell is SQuAD with Mistral-7B-Instruct
\citep{jiang2023mistral}. Faithfulness is scored with DeBERTa-v3 NLI
\citep{he2021deberta}, a second NLI model, and a local RAGAS-style
Mistral-as-judge rubric \citep{es2024ragas}. The local judge is useful for
triangulation but may share family-level biases with the generator; we treat
that as a limitation rather than a source of authority.

\paragraph{Human calibration design.}
To ground the scorer disagreement, two raters annotate 99 examples sampled
from the fixed-generation multi-metric cell. Labels are
\texttt{supported}, \texttt{partially\_supported}, and \texttt{unsupported};
adjudicated labels are mapped to \(1,0.5,0\) for rank correlations. The
instructions require raters to judge support using only the provided passages.

% POLISH FIX 2 (2026-04-29): switched from a tabularx X column to explicit
% p{} widths with locally raised \arraystretch so multi-line cells wrap cleanly
% inside their column boundaries instead of running into the next row.
\begingroup
\renewcommand{\arraystretch}{1.15}
\begin{table}[H]
\centering
\footnotesize
\caption{\method{} audit controls. Each control can be reused with other
datasets, retrievers, generators, and faithfulness metrics.}
\label{tab:controlledrag_protocol}
\begin{tabular}{p{2.6cm}p{3.6cm}p{4.4cm}p{1.8cm}}
\toprule
Control & Claim it tests & Failure mode exposed & Primary evidence \\
\midrule
Similarity-matched intervention &
CCS improves faithfulness after local relevance is fixed &
A set statistic can lose its apparent effect once mean query-passage similarity is controlled &
Table~\ref{tab:causal} \\
Multi-seed scale correction &
A pilot effect remains stable under repeated sampling &
Small effects can be seed-sensitive even when the pilot direction is plausible &
Table~\ref{tab:scaled} \\
Fixed-generation multi-metric and human-calibrated scoring &
One automatic judge identifies the intervention effect size &
The same outputs can imply different magnitudes, and human alignment can be only moderate &
Tables~\ref{tab:metrics}--\ref{tab:human_calibration} \\
Threshold-transfer matrix &
A tuned threshold is a reusable policy &
In-domain tuning can hide weak off-diagonal transfer &
Table~\ref{tab:tauflags} \\
Coherent-vs-random noise separation &
Coherence is just generic retrieval quality &
Same-topic uninformative context and random context can have different slopes &
Table~\ref{tab:noise} \\
Cost-aware baseline audit &
A method is deployment-preferable &
Dataset wins can disappear once hallucination, p99 latency, and indexing cost are included &
Fig.~\ref{fig:pareto_p99}; Table~\ref{tab:h2h} \\
\bottomrule
\end{tabular}
\end{table}
\endgroup
